### 2024-09-30 README.qmd
Files generated by PkgTemplates
PkgTemplates version: 0.7.51

---
format:
  commonmark:
    variant: -raw_html+tex_math_dollars
    wrap: none
    mermaid-format: png
crossref:
  fig-prefix: Figure
  tbl-prefix: Table
output: asis
---

# CounterfactualTraining

[![Stable](https://img.shields.io/badge/docs-stable-blue.svg)](https://JuliaTrustworthyAI.github.io/CounterfactualTraining.jl/stable/)
[![Dev](https://img.shields.io/badge/docs-dev-blue.svg)](https://JuliaTrustworthyAI.github.io/CounterfactualTraining.jl/dev/)
[![Build Status](https://github.com/JuliaTrustworthyAI/CounterfactualTraining.jl/actions/workflows/CI.yml/badge.svg?branch=main)](https://github.com/JuliaTrustworthyAI/CounterfactualTraining.jl/actions/workflows/CI.yml?query=branch%3Amain)
[![Coverage](https://codecov.io/gh/JuliaTrustworthyAI/CounterfactualTraining.jl/branch/main/graph/badge.svg)](https://codecov.io/gh/JuliaTrustworthyAI/CounterfactualTraining.jl)
[![Code Style: Blue](https://img.shields.io/badge/code%20style-blue-4495d1.svg)](https://github.com/invenia/BlueStyle)
[![Aqua](https://raw.githubusercontent.com/JuliaTesting/Aqua.jl/master/badge.svg)](https://github.com/JuliaTesting/Aqua.jl)