### 2024-10-01 paper/related_literature.tex
finetuning loss function

\section{Related Literature} \label{sec-lit}

\citet{guo2023counternet}

\citet{ross2021learning}