### 2024-12-18 paper/presenations/slides.qmd
uff

---
title: Counterfactual Training
subtitle: Update Meeting Dec 2024
author: 
  - name: "**Patrick Altmeyer**"
    url: https://www.patalt.org/
  - name: Arie van Deursen
  - name: Cynthia C. S. Liem
institute: Delft University of Technology
date: today
format: 
  tudelft-revealjs:
    theme: custom.scss
    self-contained: true
    smaller: false
    scrollable: true
    preview-links: auto
    slide-number: true
    transition: slide
    background-transition: fade
    fig-align: center
    html-math-method: mathjax
    include-in-header:
      - text: |
          <script>
          MathJax = {
            options: {
              menuOptions: {
                settings: {
                  assistiveMml: false
                }
              }
            }
          };
          </script>
crossref: 
  prp-title: RQ
  prp-prefix: RQ
classoption: "notheorems"
---

## Status

- *Code base*: In place and streamlined for reproducibility and configuration. 
- *Experiments*: Lots of work done and results largely supportive of idea.
    - Ran into problems on DelftBlue, which has set me back about 2 weeks. 
- *Paper*: Still bare-bones. 
- *ICML*: Potentially still possible to submit something, but this will be rushed and not "finished".

## Problems on Cluster

- Trying to distribute: 
    1. Models/experiments across processes.
    2. For each model/experiment distribute the counterfactual search across processes.
- Out-of-memory issues, data races, ...
- Multi-processing for models & multi-threading for counterfactual search: low CPU efficiency on DelftBlue (jobs get cancelled).

## High-Level Idea

Combine ideas from Energy-Based Models and Adversarial Training:

$$
\ell_{\text{clf}}(f_\theta(x),y) + \lambda_{\text{gen}} \ell_{\text{gen}}(x^\prime,x;\theta) + \lambda_{\text{adv}} \ell_{\text{clf}}(f_\theta(x^\prime),y)
$$

- $x^\prime$ are counterfactuals of $x_{s}\subseteq x$.
- $\ell_{\text{gen}}$ is the difference in energies between observed samples and counterfactuals.
- Counterfactuals are recycled as adversarial examples.

## Training Details

During each `EPOCH`:

1. Generate `nce` counterfactuals and distribute across mini-batches.
2. For each batch compute:
    - Classifier loss: $\ell_{\text{clf}}(f_\theta(x),y)$
    - Generator loss: $\ell_{\text{gen}}(x^\prime,x;\theta)$
    - Adversarial loss: $\lambda_{\text{adv}} \ell_{\text{clf}}(f_\theta(x^\prime),y)$
3. Backpropagate all losses and update parameters.

## Motivation and Intuition

- Instead of using SGLD to sample from $p(x|t;\theta)$, we use counterfactual generators.
- The idea is to align counterfactual explanations with observed data to induce plausibility.
- This should only work if counterfactuals are generated faithfully (favorable evidence).
- Approach can be leveraged to impicitly encode mutability and domain constraints in model.