### 2025-01-23 paper/appendix.qmd
some work on streamline Quarto usage

---
format:
  latex:
    template: bare.tex
---

## Appendix

Here's my awesome appendix.