### 2025-01-23 paper/appendix.tex

\subsection{Appendix}\label{appendix}

Here's my awesome appendix.