### 2025-01-23 paper/related.tex

\section{Related Literature}\label{sec-lit}

\subsection{Background on Counterfactual
Explanations}\label{background-on-counterfactual-explanations}

\subsection{Learning Representations}\label{learning-representations}

\begin{quote}
For example, joint-energy models
\end{quote}

\subsection{Generalization and
Robustness}\label{generalization-and-robustness}

generate counterfactual images for MNIST and ImageNet through
independent mechanisms (IM): each IM learns class-conditional input
distributions over a specific lower-dimensional, semantically meaningful
factor, such as \emph{texture}, \emph{shape} and \emph{background}. The
demonstrate that using these generated counterfactuals during classifier
training improves model robustness. Similarly, argue that
counterfactuals represent potentially useful training data in machine
learning, especially in supervised settings where inputs may be
reasonably mapped to multiple outputs. They, too, demonstrate the
augmenting the training data of image classifiers can improve
generalization.

propose an approach using counterfactuals in training that does not rely
on data augmentation: they argue that counterfactual pairs typically
already exist in training datasets. Specifically, their approach relies
on, firstly, identifying similar input samples with different
annotations and, secondly, ensuring that the gradient of the classifier
aligns with the vector between pairs of counterfactual inputs using the
cosine distance as a loss function (referred to as \emph{gradient
supervision})

\begin{quote}
This might be useful for our task as well\ldots{}
\end{quote}

In the natural language processing (NLP) domain, counterfactuals have
similarly been used to improve models through data augmentation:
\cite{wu2021polyjuice}, propose POLYJUICE, a general-purpose
counterfactual generator for language models. They demonstrate
empirically that augmenting training data through POLYJUICE
counterfactuals improves robustness in a number of NLP tasks.

\subsection{Link to Adversarial
Training}\label{link-to-adversarial-training}

propose two definitional differences between Adversarial Examples (AE)
and Counterfactual Explanations (CE): firstly, and more importantly
according to the authors, the term AE implies missclassification, which
is not the case for CE;

\begin{quote}
this might be a useful notion for use to distinguish between
adversarials and explanations during training
\end{quote}

secondly, they argue that closeness plays a more critical role in the
context of CE but confess that even counterfactuals that are not close
might be relevant explanations. show that CE and AE are equivalent under
certain conditions and derive upper bounds on the distances between
them.

\subsection{Closely Related}\label{closely-related}

are the first to propose end-to-end training pipeline that includes
counterfactual explanations as part of the training prodeduce. In
particular, they propose a specific network architecture that includes a
predictor and CE generator network,

\begin{quote}
akin a GAN?
\end{quote}

where the parameters of the CE generator network are learnable.
Counterfactuals are generated during each training iteration and fed
back to the predictor network (\textcolor{red}{here we are aligned}). In
contrast, we impose no restrictions on the neural network architecture
at all.

\begin{quote}
to ensure the one-hot encoding of categorical features is maintained,
they simple use softmax (might be interesting for CE.jl)
\end{quote}

Interestingly, the authors find that their approach is sensitive to the
choice of the loss function: only MSE seems to lead to good performance.
They also demonstrate theoretically, that the objective function is
difficult to optimize due to divergent gradients and suffers from poor
adversarial robustness.

\begin{quote}
because partial gradients with respect to the classification loss
component and the counterfactual validity component point in opposite
directions
\end{quote}

To mitigate these issues, the authors use block-wise gradient descent:
they first update with respect to classification loss and then use a
second update with respect to the other loss components

\begin{quote}
this might be useful for our task as well
\end{quote}

propose a way to train models that are guaranteed to provide recourse
for individuals with high probability. The approach builds on
adversarial training,

\begin{quote}
here we are aligned
\end{quote}

where in this context adversarial examples are actively encouraged to
exist, but only target attacks with respect to the positive class. The
proposed method allows for imposing a set of actionable recourse
ex-ante: for example, users can impose mutability constraints for
features.

\begin{quote}
here we are aligned
\end{quote}

\begin{quote}
To solve their objective function more efficiently, they use a
first-order Taylor approximation to approximate the recourse loss
component (might be applicable in our case
\end{quote}

introduce Counterfactual Adversarial Training (CAT) with intention of
improving generalization and robustness of language models.
Specifically, they propose to proceed as follows: firstly, identify
training samples that are subject to high predictive uncertainty
(entropy); secondly, generate counterfactual explanations for those
samples; and, finally, finetune the model on the augmented dataset that
includes the generated counterfactuals.