### 2025-01-24 paper/appendix.qmd
might need plots after all

```{julia}
using CTExperiments
using CTExperiments.CSV
using CTExperiments.DataFrames
using CTExperiments.StatsBase
using DotEnv

DotEnv.load!()
```

# Appendix

## Initial Grid Search

```{julia}
results_dir = ENV["INITIAL_RUN"]
```

### Generator Params

```{julia}
gen_params_dir = joinpath(results_dir, "gen_params/mlp")
```

#### Linearly Separable

```{julia}
df = CSV.read("$(gen_params_dir)/lin_sep/evaluation/results/ce/objective---lambda_energy_exper---lambda_energy_eval/plausibility_distance_from_target.csv", DataFrame)
df = groupby(df, Not(:run, :std, :mean, :lambda_energy_eval)) |> 
  gdf -> combine(gdf, :mean => (x -> [(mean(x),std(x))]) => [:value,:std]) |>
  df -> sort(df, [:lambda_energy_exper, :objective, :generator_type]) |>
  df -> subset(df, :lambda_energy_exper => x -> x .>= 0.5)
tabulate_results(df, "value"; alpha=0.9)
```

#### Moons

```{julia}
df = CSV.read("$(gen_params_dir)/moons/evaluation/results/ce/objective---lambda_energy_exper---lambda_energy_eval/plausibility_distance_from_target.csv", DataFrame)
df = groupby(df, Not(:run, :std, :mean, :lambda_energy_eval)) |> 
  gdf -> combine(gdf, :mean => (x -> [(mean(x),std(x))]) => [:value,:std]) |>
  df -> sort(df, [:lambda_energy_exper, :objective, :generator_type]) |>
  df -> subset(df, :lambda_energy_exper => x -> x .>= 0.5)
tabulate_results(df, "value"; alpha=0.9)
```

#### Circles

```{julia}
df = CSV.read("$(gen_params_dir)/circles/evaluation/results/ce/objective---lambda_energy_exper---lambda_energy_eval/plausibility_distance_from_target.csv", DataFrame)
df = groupby(df, Not(:run, :std, :mean, :lambda_energy_eval)) |> 
  gdf -> combine(gdf, :mean => (x -> [(mean(x),std(x))]) => [:value,:std]) |>
  df -> sort(df, [:lambda_energy_exper, :objective, :generator_type]) |>
  df -> subset(df, :lambda_energy_exper => x -> x .>= 0.5)
tabulate_results(df, "value"; alpha=0.9)
```