### 2025-02-14 paper/sections/other/constraints.qmd

```{julia}
#| eval: false

projectdir = splitpath(pwd()) |>
    ss -> joinpath(ss[1:findall([s == "CounterfactualTraining.jl" for s in ss])[1]]...) 
cd(projectdir)

using CTExperiments
using CTExperiments.CounterfactualExplanations
using CTExperiments.CounterfactualTraining
using CTExperiments.Flux
using CTExperiments.Plots
using CTExperiments.TaijaPlotting
using Plots.PlotMeasures
```

## Protecting Mutability Constraints with Linear Classifiers {#sec-app-constraints}

In @sec-constraints we explain that to avoid penalizing implausibility that arises due to mutability constraints, we impose a point mass prior on $p(\mathbf{x})$ for the corresponding feature. We argue in @sec-constraints that this approach induces models to be less sensitive to immutable features and demonstrate this empirically in @sec-experiments. Below we derive the analytical results in @thm-mtblty.

::: {.proof}

Let $d_{\text{mtbl}}$ and $d_{\text{immtbl}}$ denote some mutable and immutable feature, respectively. Suppose that $\mu_{y^-,d_{\text{immtbl}}} < \mu_{y^+,d_{\text{immtbl}}}$ and $\mu_{y^-,d_{\text{mtbl}}} > \mu_{y^+,d_{\text{mtbl}}}$, where $\mu_{k,d}$ denotes the conditional sample mean of feature $d$ in class $k$. In words, we assume that the immutable feature tends to take lower values for samples in the non-target class $y^-$ than in the target class $y^+$. We assume the opposite to hold for the mutable feature.

Assuming multivariate Gaussian class densities with common diagonal covariance matrix $\Sigma_k=\Sigma$ for all $k \in \mathcal{K}$, we have for the log likelihood ratio between any two classes $k,m \in \mathcal{K}$ [@hastie2009elements]:

$$
\log \frac{p(k|\mathbf{x})}{p(m|\mathbf{x})}=\mathbf{x}^\intercal \Sigma^{-1}(\mu_{k}-\mu_{m})  + \text{const}
$$ {#eq-loglike}

By independence of $x_1,...,x_D$, the full log-likelihood ratio decomposes into:

$$
\log \frac{p(k|\mathbf{x})}{p(m|\mathbf{x})} = \sum_{d=1}^D \frac{\mu_{k,d}-\mu_{m,d}}{\sigma_{d}^2} x_{d} + \text{const}
$$ {#eq-loglike-decomp}

By the properties of our classifier (*multinomial logistic regression*), we have:

$$
\log \frac{p(k|\mathbf{x})}{p(m|\mathbf{x})} = \sum_{d=1}^D \left( \theta_{k,d} - \theta_{m,d} \right)x_d + \text{const}
$$ {#eq-multi}

where $\theta_{k,d}=\Theta[k,d]$ denotes the coefficient on feature $d$ for class $k$.

Based on @eq-loglike-decomp and @eq-multi we can identify that $(\mu_{k,d}-\mu_{m,d}) \propto (\theta_{k,d} - \theta_{m,d})$ under the assumptions we made above. Hence, we have that $(\theta_{y^-,d_{\text{immtbl}}} - \theta_{y^+,d_{\text{immtbl}}}) < 0$ and $(\theta_{y^-,d_{\text{mtbl}}} - \theta_{y^+,d_{\text{mtbl}}}) > 0$

Let $\mathbf{x}^\prime$ denote some randomly chosen individual from class $y^-$ and let $y^+ \sim p(y)$ denote the randomly chosen target class. Then the partial derivative of the contrastive divergence penalty [@eq-div] with respect to coefficient $\theta_{y^+,d}$ is equal to

$$
\frac{\partial}{\partial\theta_{y^+,d}} \left(\text{div}(\mathbf{x},\mathbf{x^\prime},\mathbf{y};\theta)\right) = \frac{\partial}{\partial\theta_{y^+,d}} \left( \left(-\mathbf{M}_\theta(\mathbf{x})[y^+]\right) - \left(-\mathbf{M}_\theta(\mathbf{x}^\prime)[y^+]\right) \right) = x_{d}^\prime - x_{d}
$$ {#eq-grad}

and equal to zero everywhere else.

Since $(\mu_{y^-,d_{\text{immtbl}}} < \mu_{y^+,d_{\text{immtbl}}})$ we are more likely to have $(x_{d_{\text{immtbl}}}^\prime - x_{d_{\text{immtbl}}}) < 0$ than vice versa at initialization. Similarly, we are more likely to have $(x_{d_{\text{mtbl}}}^\prime - x_{d_{\text{mtbl}}}) > 0$ since $(\mu_{y^-,d_{\text{mtbl}}} > \mu_{y^+,d_{\text{mtbl}}})$.

This implies that if we do not protect feature $d_{\text{immtbl}}$, the contrastive divergence penalty will push down on $\theta_{y^-,d_{\text{immtbl}}}$ thereby exacerbating the existing effect $(\theta_{y^-,d_{\text{immtbl}}} - \theta_{y^+,d_{\text{immtbl}}}) < 0$. In words, not protecting the immutable feature would have the undesirable effect of making the classifier more sensitive to this feature, in that it would be more likely to predict class $y^-$ as opposed to $y^+$ for lower values of $d_{\text{immtbl}}$.

By the same rationale, the contrastive divergence penalty can generally be expected to push up on $\theta_{y^-,d_{\text{mtbl}}}$ exacerbating $(\theta_{y^-,d_{\text{mtbl}}} - \theta_{y^+,d_{\text{mtbl}}}) > 0$. In words, this has the effect of making the classifier more sensitive to the mutable feature, in that it would be more likely to predict class $y^-$ as opposed to $y^+$ for higher values of $d_{\text{mtbl}}$.

Thus, our proposed approach of protecting feature $d_{\text{immtbl}}$ has the net affect of decreasing the classifier's sensitivity to the immutable feature relative to the mutable feature (i.e. no change in sensitivity for $d_{\text{immtbl}}$ relative to increased sensitivity for $d_{\text{mtbl}}$).

:::

::: {.callout-warning}

\@Cynthia, \@Arie, I have tentatively phrased the above in terms of a theorem and proof. This is something I've so far shied away from because I feel a bit out of my depth when it comes to mathematical proofs. The above makes intuitive sense to me, but I don't know for sure if it's correct.

:::

```{julia}
#| eval: false

using CTExperiments: 
    get_ce_data, 
    train_val_split, 
    build_model, 
    LinearModel, 
    get_input_encoder

# Callback:
using CTExperiments: get_log_reg_params, get_decision_boundary

function plot_db(model, ces; _xlab="Debt", _ylab="Age")

    x = (ce -> ce.counterfactual).(ces) |> xs -> reduce(hcat, xs)
    x0 = (ce -> ce.factual).(ces) |> xs -> reduce(hcat, xs)

    xlab = _constraint[1] == "both" ? "$(_xlab) (mutable)" : "$(_xlab) (immutable)"
    ylab = _constraint[2] == "both" ? "$(_ylab) (mutable)" : "$(_ylab) (immutable)"

    # Data and decision boundary:
    coeff = get_log_reg_params(model)
    db = get_decision_boundary(coeff)
    plt = Plots.scatter(
        ce_data, 
        # label=["1=Default" "2=No Default"], 
        # legend_position=:topright, 
        xlab=xlab, ylab=ylab,  
        axis=nothing, 
        legend=false,
        title=_title
    )
    Plots.abline!(plt, db.slope, db.intercept; lw=5, label="Dec. Boundary")

    # Counterfactuals:
    if !any(isnothing.(x))
        yhat = [argmax(y) for y in eachcol(model(x))]
        yhat0 = [argmax(y) for y in eachcol(model(x0))]
        if any(yhat.==2)

            # Paths:
            u = []
            v = []
            for (i,ce) in enumerate(eachcol(x))
                Δ = ce - x0[:,i]
                push!(u, Δ[1])
                push!(v, Δ[2])
            end
            Plots.quiver!(x0[1,yhat.==2], x0[2,yhat.==2], quiver=(u[yhat.==2], v[yhat.==2]), color=1)

            # End points:
            Plots.scatter!(x[1,yhat.==2], x[2,yhat.==2], label=["CE (y⁺=1)" "CE (y⁺=2)"], ms=15, shape=:star, color=yhat[yhat.==2], group=yhat[yhat.==2], mscolor=yhat0[yhat.==2])

        end
    end
    display(plt)
end

# Data:
specs = [
    ("(a)", VanillaObjective(needs_ce=true), ["both", "both"]),
    ("(b)", EnergyDifferentialObjective(lambda=[1.0,1.0,0.0]), ["both", "both"]),
    ("(c)", VanillaObjective(needs_ce=true), ["both", "none"]),
    ("(d)", EnergyDifferentialObjective(lambda=[1.0,1.0,0.0]), ["both", "none"]),
]
```

```{julia}
#| eval: false

using CounterfactualExplanations.Convergence
using Random
Random.seed!(42)

models = []
plts = []

for (title, obj, constraint) in specs

    # Globals for the callback:
    global _title = title
    global _constraint = constraint

    # Data:
    data = LinearlySeparable(
        n_train=500,
        batchsize=50,
        mutability=constraint
    )
    global ce_data = get_ce_data(data)
    val_size = data.n_validation / (data.n_validation + data.n_train)
    train_set, val_set, _ = train_val_split(data, ce_data, val_size)

    # Model:
    nin = size(first(train_set)[1], 1)
    nout = size(first(train_set)[2], 1)
    model = build_model(LinearModel(), nin, nout)

    # Objective:
    generator = GenericGenerator()
    opt_state = Flux.setup(Descent(), model)
    conv = MaxIterConvergence()

    model, logs = counterfactual_training(
        obj,
        model,
        generator,
        train_set,
        opt_state;
        val_set = val_set,
        nepochs = 50,
        mutability = Symbol.(constraint),
        callback = plot_db,
        nce=50,
        convergence=conv, 
    )

    push!(models, model)
    push!(plts, current())
end
```

```{julia}
#| eval: false

plt = Plots.plot(
    plts..., 
    layout=(1,4), 
    size=(1150,250),  
    left_margin = 10mm, 
    bottom_margin = 5mm,
    top_margin = 3mm,
    right_margin = 10mm,
)
display(plt)
Plots.savefig(plt, "paper/figures/poc.svg")
```

```{julia}
#| eval: false

plts = []
titles = ["(a)", "(b)"]
_xlab = "Existing Debt"
_ylab = "Age"
for (i,model) in enumerate(models)
    coeff = get_log_reg_params(model)
    db = get_decision_boundary(coeff)
    xlab = constraints[i][1] == "both" ? "$(_xlab) (mutable)" : "$(_xlab) (immutable)"
    ylab = constraints[i][2] == "both" ? "$(_ylab) (mutable)" : "$(_ylab) (immutable)"
    plt = Plots.scatter(ce_data; xlab=xlab, ylab=ylab, bottom_margin = 5mm, left_margin = 5mm, label=["Default" "No Default"], title=titles[i])
    Plots.abline!(plt, db.slope, db.intercept; lw=5, label="Dec. Boundary")

    push!(plts, plt)
end

plt = Plots.plot(plts..., layout=(1,2), size=(600,300))
Plots.savefig(plt, "paper/figures/app_mtblty.svg")
```