### 2025-02-19 paper/sections/other/counterfactuals.qmd
noice

## Generating Counterfactuals through Gradient Descent {#sec-app-ce}

In this section, we provide some background on gradient-based counterfactual generators (@sec-app-ce-background) and discuss how we define convergence in this context (@sec-app-conv).

### Background {#sec-app-ce-background}

### Convergence {#sec-app-conv}

An important consideration when generating counterfactual explanations using gradient-based methods is how to define convergence.