### 2025-02-19 paper/sections/other/counterfactuals.qmd

## Generating Counterfactuals through Gradient Descent {#sec-app-ce}

In this section, we provide some background on gradient-based counterfactual generators (@sec-app-ce-background) and discuss how we define convergence in this context (@sec-app-conv).

### Background {#sec-app-ce-background}

### Convergence {#sec-app-conv}

An important consideration when generating counterfactual explanations using gradient-based methods is how to define convergence. Two common choices are to 1) perform gradient descent over a fixed number of iterations $T$, or 2) conclude the search as soon as the predicted probability for the target class has reached a pre-determined threshold, $\tau$: $\mathcal{S}(\mathbf{M}_\theta(\mathbf{x}^\prime))[y^+] \geq \tau$. We prefer the latter for our purposes, because it explicitly defines convergence in terms of the black-box model, $\mathbf{M}(\mathbf{x})$.

Defining convergence in this way allows for a more intuitive interpretation of the resulting counterfactual outcomes than with fixed $T$. Specifically, it allows us to think of counterfactuals as explaining 'high-confidence' predictions by the model for the target class $y^+$. Depending on the context and application, different choices of $\tau$ can be considered as representing 'high-confidence' predictions.