### 2025-02-24 paper/sections/other/counterfactuals.qmd

## Generating Counterfactuals through Gradient Descent {#sec-app-ce}

In this section, we provide some background on gradient-based counterfactual generators (@sec-app-ce-background) and discuss how we define convergence in this context (@sec-app-conv).

### Background {#sec-app-ce-background}

Gradient-based counterfactual search was originally proposed by @wachter2017counterfactual. It generally solves the following unconstrained objective,

$$
\begin{aligned}
\min_{\mathbf{z}^\prime \in \mathcal{Z}^L} \left\{  {\text{yloss}(\mathbf{M}_{\theta}(g(\mathbf{z}^\prime)),\mathbf{y}^+)}+ \lambda {\text{cost}(g(\mathbf{z}^\prime)) }  \right\} 
\end{aligned} 
$$

where $g: \mathcal{Z} \mapsto \mathcal{X}$ is an invertible function that maps from the $L$-dimensional counterfactual state space to the feature space and $\text{cost}(\cdot)$ denotes one or more penalties that are used to induce certain properties of the counterfactual outcome. As above, $\mathbf{y}^+$ denotes the target output and $\mathbf{M}_{\theta}(\mathbf{x})$ returns the logit predictions of the underlying classifier for $\mathbf{x}=g(\mathbf{z})$.

For all generators used in this work we use standard logit crossentropy loss for $\text{yloss}(\cdot)$. All generators also penalize the distance ($\ell_1$-norm) of counterfactuals from their original factual state. For *Generic* and *ECCo*, we have $\mathcal{Z}:=\mathcal{X}$ and $g(\mathbf{z})=g(\mathbf{z})^{-1}=\mathbf{z}$, that is counterfactual are searched directly in the feature space. Conversely, *REVISE* traverses the latent space of a variational autoencoder (VAE) fitted to the training data, where $g(\cdot)$ corresponds to the decoder [@joshi2019realistic]. In addition to the distance penalty, *ECCo* uses an additional penalty component that regularizes the energy associated with the counterfactual, $\mathbf{x}^\prime$ [@altmeyer2024faithful].

### Convergence {#sec-app-conv}

An important consideration when generating counterfactual explanations using gradient-based methods is how to define convergence. Two common choices are to 1) perform gradient descent over a fixed number of iterations $T$, or 2) conclude the search as soon as the predicted probability for the target class has reached a pre-determined threshold, $\tau$: $\mathcal{S}(\mathbf{M}_\theta(\mathbf{x}^\prime))[y^+] \geq \tau$. We prefer the latter for our purposes, because it explicitly defines convergence in terms of the black-box model, $\mathbf{M}(\mathbf{x})$.

Defining convergence in this way allows for a more intuitive interpretation of the resulting counterfactual outcomes than with fixed $T$. Specifically, it allows us to think of counterfactuals as explaining 'high-confidence' predictions by the model for the target class $y^+$. Depending on the context and application, different choices of $\tau$ can be considered as representing 'high-confidence' predictions.