### 2025-02-24 paper/sections/other/compute.qmd
buff

# Computation Details {.appendix}

<!-- [@DHPC2022] -->
## Hardware

We performed our experiments on a high-performance cluster. Details about the cluster will be disclosed upon publication to avoid revealing information that might interfere with the double-blind review process. Since our experiments involve highly parallel tasks and rather small models by today's standard, we have relied on distributed computing across multiple central processing units (CPU). Graphical processing units (GPU) were not required. For larger grid searches we used up to but typically less than 100 CPUs.

## Software

All computations were performed in the Julia Programming Language [@bezanson2017julia]. We have developed a package for counterfactual training that leverages and extends the functionality provided by several existing packages, most notably [CounterfactualExplanations.jl](https://github.com/JuliaTrustworthyAI/CounterfactualExplanations.jl) [@altmeyer2023explaining] and the [Flux.jl](https://fluxml.ai/Flux.jl/v0.16/) library for deep learning [@innes2018fashionable;@innes2018flux]. For data-wrangling and presentation-ready tables we relied on [DataFrames.jl](https://dataframes.juliadata.org/v1.7/) [@milan2023dataframes] and [PrettyTables.jl](https://ronisbr.github.io/PrettyTables.jl/v2.4/) [@chagas2024pretty], respectively. For plots and visualizations we used both [Plots.jl](https://docs.juliaplots.org/v1.40/) [@PlotsJL] and [Makie.jl](https://docs.makie.org/v0.22/) [@danisch2021makie], in particular [AlgebraOfGraphics.jl](https://aog.makie.org/v0.9.3/). To distribute computational tasks across multiple processors, we have relied on [MPI.jl](https://juliaparallel.org/MPI.jl/v0.20/) [@byrne2021mpi].