### 2025-03-06 paper/sections/appendix.qmd
more work on appendix

\FloatBarrier

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

<!-- # Supplementary Material {.appendix} -->

{{< include /paper/sections/other/notation.qmd >}}

# Technical Details of Our Approach {.appendix}

{{< include /paper/sections/other/counterfactuals.qmd >}}

{{< include /paper/sections/other/constraints.qmd >}}

{{< include /paper/sections/other/training.qmd >}}

# Detailed Results {.appendix}

## Qualitative Findings for Image Data

::: {.callout-note}

@fig-mnist shows much more plausible (faithful) counterfactuals for a model with CT than the model with conventional training (@fig-mnist-vanilla). In fact, this is not even using *ECCo+* and still showing better results than the best results we achieved in our AAAI paper for JEM ensembles.

:::

![Counterfactual images for *MLP* with counterfactual training. The underlying generator, *ECCo*, aims to generate counterfactuals that are faithful to the model [@altmeyer2024faithful].](/paper/figures/mnist_mlp.png){#fig-mnist}

![Counterfactual images for *MLP* with conventional training. The underlying generator, *ECCo*, aims to generate counterfactuals that are faithful to the model [@altmeyer2024faithful].](/paper/figures/mnist_mlp_vanilla.png){#fig-mnist-vanilla}

{{< include /paper/sections/other/grids.qmd >}}

<!-- {{< include /paper/sections/other/initial_grids.qmd >}} -->

{{< include /paper/sections/other/compute.qmd >}}