### 2025-03-07 paper/sections/other/grids.qmd
some notes from meeting and updates to appendix

## Grid Searches {#sec-app-grid}

```{julia}
projectdir = splitpath(pwd()) |>
    ss -> joinpath(ss[1:findall([s == "CounterfactualTraining.jl" for s in ss])[1]]...) 
cd(projectdir)

using CTExperiments
using CTExperiments.CSV
using CTExperiments.DataFrames
using CTExperiments.StatsBase

using DotEnv
DotEnv.load!()
```

```{julia}
res_dir = ENV["FINAL_GRID_RESULTS"]
```

### Generator Parameters

```{julia}
grid_dir = joinpath(res_dir, "gen_params/mlp")
data_dirs = readdir(grid_dir) |> x -> joinpath.(grid_dir, x) |> x -> x[isdir.(x)]
data_names = basename.(data_dirs)
suffix = "evaluation/results/ce/decision_threshold_exper---lambda_energy_exper---maxiter_exper---maxiter---decision_threshold_exper/"
```

The hyperparameter grid with varying generator parameters during training is shown in @nte-gen-params-final-run-train. The corresponding evaluation grid used for these experiments is shown in @nte-gen-params-final-run-eval.

::: {#nte-gen-params-final-run-train .callout-note}

## Training Phase

```{julia}
#| output: asis
dict = CTExperiments.from_toml(joinpath(grid_dir, "lin_sep/grid_config.toml")) 
println(CTExperiments.dict_to_quarto_markdown(dict))
```

:::

::: {#nte-gen-params-final-run-eval .callout-note}

## Evaluation Phase

```{julia}
#| output: asis
dict = CTExperiments.from_toml(joinpath(grid_dir, "lin_sep/evaluation/evaluation_grid_config.toml"))
println(CTExperiments.dict_to_quarto_markdown(dict))
```

:::

#### Plausibility

```{julia}
#| output: asis

fig_label_prefix = "grid-gen_params-plaus"
fig_labels = (nm -> "fig-$(fig_label_prefix)-$nm").(data_names)
_str = "The results with respect to the plausibility measure are shown in @$(fig_labels[1]) to @$(fig_labels[end])."
println(_str)
```

```{julia}
#| output: asis

imgfname = "plausibility_distance_from_target.png"
fig_caption = "Average outcomes for the plausibility measure across hyperparameters."
full_paths = joinpath.(data_dirs, joinpath(suffix,imgfname))
include_img_commands = CTExperiments.get_img_command(data_names, full_paths, fig_labels; fig_caption) 
_str = join(include_img_commands, "\n\n")
println(_str)
```

#### Cost

```{julia}
#| output: asis

fig_label_prefix = "grid-gen_params-cost"
fig_labels = (nm -> "fig-$(fig_label_prefix)-$nm").(data_names)
_str = "The results with respect to the cost measure are shown in @$(fig_labels[1]) to @$(fig_labels[end])."
println(_str)
```

```{julia}
#| output: asis

imgfname = "distance.png"
fig_caption = "Average outcomes for the cost measure across hyperparameters."
full_paths = joinpath.(data_dirs, joinpath(suffix,imgfname))
include_img_commands = CTExperiments.get_img_command(data_names, full_paths, fig_labels; fig_caption)
_str = join(include_img_commands, "\n\n")
println(_str)
```

### Penalty Strengths

```{julia}
grid_dir = joinpath(res_dir, "penalties/mlp")
data_dirs = readdir(grid_dir) |> x -> joinpath.(grid_dir, x) |> x -> x[isdir.(x)]
data_names = basename.(data_dirs)
suffix = "evaluation/results/ce/lambda_adversarial---lambda_energy_reg---lambda_energy_diff---lambda_adversarial---lambda_adversarial/"
```

The hyperparameter grid with varying penalty strengths during training is shown in @nte-pen-final-run-train. The corresponding evaluation grid used for these experiments is shown in @nte-pen-final-run-eval.

::: {#nte-pen-final-run-train .callout-note}

## Training Phase

```{julia}
#| output: asis
dict = CTExperiments.from_toml(joinpath(grid_dir, "lin_sep/grid_config.toml")) 
println(CTExperiments.dict_to_quarto_markdown(dict))
```

:::

::: {#nte-pen-final-run-eval .callout-note}

## Evaluation Phase

```{julia}
#| output: asis
dict = CTExperiments.from_toml(joinpath(grid_dir, "lin_sep/evaluation/evaluation_grid_config.toml"))
println(CTExperiments.dict_to_quarto_markdown(dict))
```

:::

#### Plausibility

```{julia}
#| output: asis

fig_label_prefix = "grid-pen-plaus"
fig_labels = (nm -> "fig-$(fig_label_prefix)-$nm").(data_names)
_str = "The results with respect to the plausibility measure are shown in @$(fig_labels[1]) to @$(fig_labels[end])."
println(_str)
```

```{julia}
#| output: asis

imgfname = "plausibility_distance_from_target.png"
fig_caption = "Average outcomes for the plausibility measure across hyperparameters."
full_paths = joinpath.(data_dirs, joinpath(suffix,imgfname))
include_img_commands = CTExperiments.get_img_command(data_names, full_paths, fig_labels; fig_caption) 
_str = join(include_img_commands, "\n\n")
println(_str)
```

#### Cost

```{julia}
#| output: asis

fig_label_prefix = "grid-pen-cost"
fig_labels = (nm -> "fig-$(fig_label_prefix)-$nm").(data_names)
_str = "The results with respect to the cost measure are shown in @$(fig_labels[1]) to @$(fig_labels[end])."
println(_str)
```

```{julia}
#| output: asis

imgfname = "distance.png"
fig_caption = "Average outcomes for the cost measure across hyperparameters."
full_paths = joinpath.(data_dirs, joinpath(suffix,imgfname))
include_img_commands = CTExperiments.get_img_command(data_names, full_paths, fig_labels; fig_caption)
_str = join(include_img_commands, "\n\n")
println(_str)
```