### 2025-03-10 paper/sections/other/compute.qmd
present tunign results

# Computation Details {.appendix}

<!-- [@DHPC2022] -->
## Hardware

We performed our experiments on a high-performance cluster. Details about the cluster will be disclosed upon publication to avoid revealing information that might interfere with the double-blind review process. Since our experiments involve highly parallel tasks and rather small models by today's standard, we have relied on distributed computing across multiple central processing units (CPU). Graphical processing units (GPU) were not required.

### Grid Searches

Model training for the largest grid searches with 270 unique parameter combinations was parallelized across 34 CPUs with 2GB memory each. The time to completion varied by dataset for reasons discussed in @sec-discussion: 0h49m (*Moons*), 1h4m (*Linearly Separable*), 1h49m (*Circles*), 3h52m (*Overlapping*). Model evaluations for large grid searches were parallelized across 20 CPUs with 3GB memory each. Evaluations for all data sets took less than one hour (<1h) to complete.

### Tuning

For tuning of selected hyperparameters, we distributed the task of generating counterfactuals during training across 40 CPUs with 2GB memory each for all tabular datasets. Except for the *Adult* dataset, all training runs were completed in less that half an hour (<0h30m). The *Adult* dataset took around 0h35m to complete. For *MNIST*, we relied on 100 CPUs with 2GB memory each.

## Software

All computations were performed in the Julia Programming Language [@bezanson2017julia]. We have developed a package for counterfactual training that leverages and extends the functionality provided by several existing packages, most notably [CounterfactualExplanations.jl](https://github.com/JuliaTrustworthyAI/CounterfactualExplanations.jl) [@altmeyer2023explaining] and the [Flux.jl](https://fluxml.ai/Flux.jl/v0.16/) library for deep learning [@innes2018fashionable;@innes2018flux]. For data-wrangling and presentation-ready tables we relied on [DataFrames.jl](https://dataframes.juliadata.org/v1.7/) [@milan2023dataframes] and [PrettyTables.jl](https://ronisbr.github.io/PrettyTables.jl/v2.4/) [@chagas2024pretty], respectively. For plots and visualizations we used both [Plots.jl](https://docs.juliaplots.org/v1.40/) [@PlotsJL] and [Makie.jl](https://docs.makie.org/v0.22/) [@danisch2021makie], in particular [AlgebraOfGraphics.jl](https://aog.makie.org/v0.9.3/). To distribute computational tasks across multiple processors, we have relied on [MPI.jl](https://juliaparallel.org/MPI.jl/v0.20/) [@byrne2021mpi].