### 2025-03-10 paper/sections/other/tuning.qmd

## Hyperparameter Tuning {#sec-app-tune}

```{julia}
projectdir = splitpath(pwd()) |>
    ss -> joinpath(ss[1:findall([s == "CounterfactualTraining.jl" for s in ss])[1]]...) 
cd(projectdir)

using CTExperiments
using CTExperiments.CSV
using CTExperiments.DataFrames
using CTExperiments.StatsBase

using DotEnv
DotEnv.load!()
```

```{julia}
res_dir = ENV["FINAL_GRID_RESULTS"]
```

Based on the findings from our initial large grid searches (@sec-app-grid), we tune selected hyperparameters for all datasets.