### 2025-03-12 paper/sections/other/tuning.qmd

## Hyperparameter Tuning {#sec-app-tune}

```{julia}
projectdir = splitpath(pwd()) |>
    ss -> joinpath(ss[1:findall([s == "CounterfactualTraining.jl" for s in ss])[1]]...) 
cd(projectdir)

using CTExperiments
using CTExperiments.CSV
using CTExperiments.DataFrames
using CTExperiments.StatsBase

using DotEnv
DotEnv.load!()
```

```{julia}
res_dir = ENV["FINAL_GRID_RESULTS"]
```

Based on the findings from our initial large grid searches (@sec-app-grid), we tune selected hyperparameters for all datasets: namely, the decision threshold $\tau$ and the strength of the energy regularization $\lambda_{\text{reg}}$. The final hyperparameter choices for each dataset are presented in **ADD TABLE**. Detailed results for each data set are shown in **ADD FIGURES**. From **ADD TABLE**, we notice that the same decision threshold of $\tau=0.5$ is optimal for all but on dataset. We attribute this to the fact that a low decision threshold results in a higher share of mature counterfactuals and hence more opportunities for the model to learn from examples. This has played a role in particular for our real-world tabular datasets and MNIST, which suffered from low levels of maturity for higher decision thresholds. In cases where maturity is not an issue, as for *Moons*, higher decision thresholds lead to better outcomes, which may have to do with the fact that the resulting counterfactuals are more faithful to the model. Concerning the regularization strength, we find somewhat high variation across datasets. Most notably, we find that relatively low levels of regularization are optimal for MNIST. We hypothesize that this finding may be attributed to the uniform scaling of all input features (digits). Finally, to increase the proportion of mature counterfactuals for some datasets, we have also investigated the effect on the learning rate $\eta$ for the counterfactual search, but found little effect on the results.