### 2025-03-13 paper/sections/appendix.qmd

\FloatBarrier

<!-- \setcounter{section}{0} -->
\renewcommand{\thesection}{\Alph{section}}

<!-- \setcounter{table}{0} -->
\renewcommand{\thetable}{A\arabic{table}}

<!-- \setcounter{figure}{0} -->
\renewcommand{\thefigure}{A\arabic{figure}}

<!-- # Supplementary Material {.appendix} -->

{{< include /paper/sections/other/notation.qmd >}}

{{< include /paper/sections/other/technical.qmd >}}

{{< include /paper/sections/other/main_results.qmd >}}

{{< include /paper/sections/other/grids.qmd >}}

{{< include /paper/sections/other/tuning.qmd >}}

{{< include /paper/sections/other/compute.qmd >}}