### 2025-03-15 paper/sections/other/training.qmd

```{julia}
projectdir = splitpath(pwd()) |>
    ss -> joinpath(ss[1:findall([s == "CounterfactualTraining.jl" for s in ss])[1]]...) 
cd(projectdir)

using CTExperiments
using CTExperiments.CSV
using CTExperiments.DataFrames
using CTExperiments.StatsBase
using Logging

using DotEnv
DotEnv.load!()
```

## Training Details {#sec-app-training}

In this section, we describe the training procedure in detail. While the details laid out here are not crucial for understanding our proposed approach, they are of importance to anyone looking to implement counterfactual training.

### Default Hyperparameters

@nte-train-default presents the default hyperparameters used during training.

::: {#nte-train-default .callout-note}

## Training Phase

```{julia}
#| output: asis

dict = Logging.with_logger(Logging.NullLogger()) do 
    CTExperiments.to_dict(Experiment()) 
end
println(CTExperiments.dict_to_quarto_markdown(dict, filter_empty=false))
```

:::

## Evaluation Details {#sec-app-eval}

@nte-eval-default presents the default hyperparameters used during evaluation.

::: {#nte-eval-default .callout-note}

## Evaluation Phase

```{julia}
#| output: asis

dict = Logging.with_logger(Logging.NullLogger()) do  
    CTExperiments.to_dict(EvaluationConfig(ExperimentGrid()))
end
println(CTExperiments.dict_to_quarto_markdown(dict, filter_empty=false))
```

:::