### 2025-03-17 paper/sections/appendix.qmd

\FloatBarrier

<!-- \renewcommand{\thesection}{\Alph{section}}

\renewcommand{\thetable}{A\arabic{table}}

\renewcommand{\thefigure}{A\arabic{figure}} -->

{{< include /paper/sections/other/notation.qmd >}}

{{< include /paper/sections/other/technical.qmd >}}

{{< include /paper/sections/other/main_results.qmd >}}

\FloatBarrier

{{< include /paper/sections/other/grids.qmd >}}

\FloatBarrier

{{< include /paper/sections/other/tuning.qmd >}}

\FloatBarrier

{{< include /paper/sections/other/compute.qmd >}}