### 2025-07-17 paper/quarto_aaai/paper.tex
removed figure and synthetic example from experiments

\makeatletter
\@ifpackageloaded{hyperref}{}{%
  \newcommand{\texorpdfstring}[2]{#1}%
}
\newcommand{\tightlist}
\makeatother
\providecommand\phantomsection{}

%File: anonymous-submission-latex-2026.tex
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage[submission]{aaai2026}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet}  % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in} % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in} % DO NOT CHANGE THIS
%
% These are recommended to typeset algorithms but not required. See the subsubsection on algorithms. Remove them if you don't have algorithms in your paper.
\usepackage{algorithm}
\usepackage{algorithmic}

%
% These are are recommended to typeset listings but not required. See the subsubsection on listing. Remove this block if you don't have listings in your paper.
\usepackage{newfloat}
\usepackage{listings}
\DeclareCaptionStyle{ruled}{labelfont=normalfont,labelsep=colon,strut=off} % DO NOT CHANGE THIS
\lstset{%
	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
	aboveskip=0pt,belowskip=0pt,%
	showstringspaces=false,tabsize=2,breaklines=true}
\floatstyle{ruled}
\newfloat{listing}{tb}{lst}{}
\floatname{listing}{Listing}
%
% Keep the \pdfinfo as shown here. There's no need
% for you to add the /Title and /Author tags.
\pdfinfo{
/TemplateVersion (2026.1)
}

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{2} %May be changed to 1 or 2 if section numbers are desired.

% The file aaai2026.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%

% Title

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
% \title{AAAI Press Anonymous Submission\\Instructions for Authors Using \LaTeX{}}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

    % Sunil Issar\textsuperscript{\rm 2},
    % J. Scott Penberthy\textsuperscript{\rm 3},
    % George Ferguson\textsuperscript{\rm 4},
    % Hans Guesgen\textsuperscript{\rm 5}
    % Note that the comma should be placed after the superscript

    1101 Pennsylvania Ave, NW Suite 300\\
    Washington, DC 20004 USA\\
    % email address must be in roman text type, not monospace or sans serif
    proceedings-questions@aaai.org
%
% See more examples next
}

%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
\iffalse
\title{My Publication Title --- Single Author}
\author {
    Author Name
}
\affiliations{
    Affiliation\\
    Affiliation Line 2\\
    name@example.com
}
\fi

\iffalse
%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
% \title{My Publication Title --- Multiple Authors}
\author {
    % Authors
    First Author Name\textsuperscript{\rm 1},
    Second Author Name\textsuperscript{\rm 2},
    Third Author Name\textsuperscript{\rm 1}
}
\affiliations {
    % Affiliations
    \textsuperscript{\rm 1}Affiliation 1\\
    \textsuperscript{\rm 2}Affiliation 2\\
    firstAuthor@affiliation1.com, secondAuthor@affilation2.com, thirdAuthor@affiliation1.com
}
\fi
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\usepackage{placeins}
\usepackage{siunitx}
\sisetup{uncertainty-mode = separate}

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\title{Counterfactual Training: Teaching Models Plausible and Actionable
Explanations}
\author{Patrick Altmeyer \and Aleksander Buszydlik \and Arie van
Deursen \and Cynthia C. S. Liem}
\date{}
\begin{document}
\maketitle
\begin{abstract}
We propose a novel training regime termed counterfactual training that
leverages counterfactual explanations to increase the explanatory
capacity of models. Counterfactual explanations have emerged as a
popular post-hoc explanation method for opaque machine learning models:
they inform how factual inputs would need to change in order for a model
to produce some desired output. To be useful in real-word
decision-making systems, counterfactuals should be (1) plausible with
respect to the underlying data and (2) actionable with respect to the
user-defined mutability constraints. Much existing research has
therefore focused on developing post-hoc methods to generate
counterfactuals that meet these desiderata. In this work, we instead
hold models directly accountable for the desired end goal:
counterfactual training employs counterfactuals ad-hoc during the
training phase to minimize the divergence between learned
representations and plausible, actionable explanations. We demonstrate
empirically and theoretically that our proposed method facilitates
training models that deliver inherently desirable explanations while
promoting robustness and preserving high predictive performance.
\end{abstract}
\section{Introduction}\label{introduction}

Today's prominence of artificial intelligence (AI) has largely been
driven by the success of representation learning with large degrees of
freedom: instead of relying on features and rules hand-crafted by
humans, modern machine learning (ML) models are tasked with learning
highly complex representations directly from the data, guided by narrow
objectives such as predictive accuracy \citep{goodfellow2016deep}. These
models tend to be so complex and opaque that humans cannot easily
interpret their decision-making.

Counterfactual explanations (CE) have become an integral part of the
broader explainable AI (XAI) toolkit \citep{molnar2022interpretable}
that can be employed to make sense of this complexity. Originally
proposed by \citet{wachter2017counterfactual}, CEs prescribe minimal
changes for factual inputs that, if implemented, would prompt some
fitted model to produce an alternative, more desirable output
(Figure~\ref{fig-poc}). This is useful and necessary to not only
understand how opaque models make predictions, but also to provide
recourse to individuals subjected to them: a retail bank, for example,
could use CE to provide meaningful feedback to unsuccessful loan
applicants that were rejected based on an opaque AI decision-making
system (ADM).

For feedback to be meaningful, counterfactual explanations need to
fulfill certain desiderata
\citep{verma2020counterfactual, karimi2020survey}: they should be
faithful to the model \citep{altmeyer2024faithful}, plausible
\citep{joshi2019realistic} and actionable \citep{ustun2019actionable}.
Plausibility is typically understood as counterfactuals being
\emph{in-domain}: unsuccessful loan applicants that implement recourse
should end up with a credit profile that is genuinely similar to that of
individuals who have successfully repaid their loans in the past.
Actionable explanations comply with practical constraints: a young,
unsuccessful loan applicant cannot increase their age in an instance.

Existing state-of-the-art (SOTA) approaches in the field have largely
focused on designing model-agnostic CE methods that identify subsets of
counterfactuals, which comply with specific desiderata. This is
problematic, because a narrow focus on any specific desideratum can
adversely affect others: it is possible, for example, to generate
plausible counterfactuals for models that are also highly vulnerable to
implausible, possibly adversarial counterfactuals
\citep{altmeyer2024faithful}. In this work, we therefore follow the
paradigm that models---as opposed to explanation methods---should be
held accountable for plausible and actionable explanations. While
previous work has shown that at least plausibility can be indirectly
achieved through existing techniques aimed at models' generative
capacity, generalization and robustness
\citep{altmeyer2024faithful, augustin2020adversarial, schut2021generating},
we directly incorporate both plausibility and actionability in the
training objective of models to improve their overall explanatory
capacity.

Specifically, \textbf{we propose Counterfactual Training (CT)}: a novel
training regime that leverages counterfactual explanations on-the-fly to
ensure that differentiable models learn plausible and actionable
explanations for the underlying data, while at the same time also being
more robust to adversarial examples (AE). Figure~\ref{fig-poc}
illustrates the outcome of CT compared to conventionally trained models:
in panel (a), faithful and valid counterfactuals end up near the
decision boundary forming a clearly distinguishable cluster in the
target class (orange); in panel (b), CT is applied to the same
underlying linear classifier architecture resulting in much more
plausible counterfactuals; in panel (c), the classifier is again trained
conventionally and we have introduced a mutability constraint on the
\emph{age} feature at test time---counterfactuals are valid but the
classifier is roughly equally sensitive to both features; by contrast,
the decision boundary in panel (c) has titled, making the model trained
with CT relatively less sensitive to the immutable \emph{age} feature.
To achieve these outcomes, CT draws inspiration from the literature on
contrastive and robust learning: we contrast faithful CE with
ground-truth data while protecting immutable features, and capitalize on
methodological links between CE and AE by penalizing the model's
adversarial loss on interim (\emph{nascent}) counterfactuals. To the
best of our knowledge, CT represents the first venture in this direction
with promising results, both empirical and theoretical.

The remainder of this manuscript is structured as follows.
Section~\ref{sec-lit} presents related work, focusing on the links to
contrastive and robust learning. Then follow our two principal
contributions. In Section~\ref{sec-method}, we introduce our
methodological framework and show theoretically that it can be employed
to enforce global actionability constraints. In our experiments
(Section~\ref{sec-experiments}), we find that thanks to counterfactual
training, (1) implausibility of CE decreases by 16\% on average; (2) the
cost of reaching valid counterfactuals with protected features decreases
by 15\%; and (3) robust accuracy increases for all datasets by up to 42
percentage points. Finally, in Section~\ref{sec-discussion}, we discuss
open challenges and conclude.

\begin{figure*}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/mediabag/../../paper/figures/poc.pdf}}

}

\caption{\label{fig-poc}Counterfactual explanations (stars) for linear
classifiers trained under different regimes on synthetic data: (a)
conventional training, all mutable; (b) CT, all mutable; (c)
conventional, \emph{age} immutable; (d) CT, \emph{age} immutable. The
linear decision boundary is shown in green along with training data
colored according to ground-truth labels: \(y^-=1\) (blue) and \(y^+=2\)
(orange). Stars indicate counterfactuals in the target class. Synthetic
features are annotated as \emph{debt} and \emph{age} for illustrative
purposes.}

\end{figure*}%

\section{Related Literature}\label{sec-lit}

To make the desiderata for our framework more concrete, we follow
previous work in tying the explanatory capacity of models to the quality
of CEs that can be generated for them
\citep{altmeyer2024faithful, augustin2020adversarial}. For simplicity,
we refer to `explanatory capacity' simply as `explainability' in what
follows (see Def. \ref{def-explainability} below for a more formal
exposition).

\subsection{Explainability and Contrastive
Learning}\label{explainability-and-contrastive-learning}

In a closely related work, \citet{altmeyer2024faithful} show that model
averaging and, in particular, contrastive model objectives can produce
more explainable and hence trustworthy models. To do so, the authors
first propose a way to generate counterfactuals that are maximally
faithful to the model in that they are consistent with what the model
has learned about the underlying data. Formally, they rely on tools from
energy-based modelling \citep{teh2003energy} to minimize the divergence
between the distribution of counterfactuals and the conditional
posterior over inputs learned by the model. Their counterfactual
explainer, \emph{ECCCo}, yields plausible explanations if and only if
the underlying model has learned representations that align with them.
The authors find that both deep ensembles
\citep{lakshminarayanan2016simple} and joint energy-based models (JEMs)
\citep{grathwohl2020your}---a form of constrastive learning---tend to do
well in this regard.

It helps to look at these findings through the lens of representation
learning with high degrees of freedom. Deep ensembles are approximate
Bayesian model averages, which are particularly effective when models
are underspecified by the available data \citep{wilson2020case}.
Averaging across solutions mitigates the aforementioned risk of
overrelying on a single locally optimal representation that corresponds
to semantically meaningless explanations for the data. Likewise,
previous work of \citet{schut2021generating} found that generating
plausible (``interpretable'') CEs is almost trivial for deep ensembles
that have undergone adversarial training. The case for JEMs is even
clearer: they optimize a hybrid objective that induces both high
predictive performance and strong generative capacity
\citep{grathwohl2020your}, which bears resemblance to the idea of
aligning models with plausible explanations and has inspired our CT
objective.

\subsection{Explainability and Robust
Learning}\label{explainability-and-robust-learning}

\citet{augustin2020adversarial} show that CEs tend to be more meaningful
(``explainable'') if the underlying model is more robust to adversarial
examples. Once again, we can make intuitive sense of this finding if we
look at adversarial training (AT) through the lens of representation
learning with high degrees of freedom: highly complex and flexible
models may learn representations that make them sensitive to implausible
or even adversarial examples \citep{szegedy2013intriguing}. Thus, by
inducing models to ``unlearn'' susceptibility to such examples,
adversarial training can effectively remove implausible explanations
from the solution space.

This interpretation of the link between explainability through
counterfactuals on the one side, and robustness to adversarial examples
on the other is backed by empirical evidence.
\citet{sauer2021counterfactual} demonstrate that using counterfactual
images during classifier training improves model robustness. Similarly,
\citet{abbasnejad2020counterfactual} argue that counterfactuals
represent potentially useful training data in machine learning,
especially in supervised settings where inputs may be reasonably mapped
to multiple outputs. They, too, show that augmenting the training data
of (image) classifiers can improve generalization performance. Finally,
\citet{teney2020learning} argue that counterfactual pairs tend to exist
in training data. Hence, their approach aims to identify similar input
samples with different annotations and ensure that the gradient of the
classifier aligns with the vector between such pairs of counterfactual
inputs using a cosine distance loss function.

CEs have also been used to improve models in the natural language
processing domain. For example, \citet{wu2021polyjuice2} propose
\emph{Polyjuice}, a general-purpose CE generator for language models and
demonstrate that the augmentation of training data with \emph{Polyjuice}
improves robustness in a number of tasks, while
\citet{luu2023counterfactual} introduce the \emph{Counterfactual
Adversarial Training} (CAT) framework that aims to improve
generalization and robustness of language models by generating
counterfactuals for training samples that are subject to high predictive
uncertainty.

There have also been several attempts at formalizing the relationship
between counterfactual explanations and adversarial examples. Pointing
to clear similarities in how CEs and AEs are generated,
\citet{freiesleben2022intriguing} makes the case for jointly studying
the opaqueness and robustness problems in representation learning.
Formally, AEs can be seen as the subset of CEs for which
misclassification is achieved \citep{freiesleben2022intriguing}.
Similarly, \citet{pawelczyk2022exploring} show that CEs and AEs are
equivalent under certain conditions.

Two other works are closely related to ours in that they use
counterfactuals during training with the explicit goal of affecting
certain properties of the post-hoc counterfactual explanations. Firstly,
\citet{ross2021learning} propose a way to train models that guarantee
recourse to a positive target class with high probability. Their
approach builds on adversarial training by explicitly inducing
susceptibility to targeted AEs for the positive class. Additionally, the
method allows for imposing a set of actionability constraints ex-ante.
For example, users can specify that certain features are immutable.
Secondly, \citet{guo2023counternet} are the first to propose an
end-to-end training pipeline that includes CEs as part of the training
procedure. Their \emph{CounterNet} network architecture includes a
predictor and a CE generator, where the parameters of the CE generator
are learnable. Counterfactuals are generated during each training
iteration and fed back to the predictor. In contrast, we impose no
restrictions on the ANN architecture at all.

\section{Counterfactual Training}\label{sec-method}

This section introduces the Counterfactual Training framework. CT
combines ideas from adversarial training, counterfactual explanations,
and energy-based modelling with the explicit goal of producing models
whose learned representations align with plausible explanations that
further comply with user-defined actionability constraints.

In the context of counterfactual explanations, plausibility has broadly
been defined as the degree to which generated CEs comply with the
underlying data-generating process
\citep{altmeyer2024faithful, guidotti2022counterfactual, poyiadzi2020face}.
Plausibility is a necessary but insufficient condition for using CEs to
provide algorithmic recourse (AR) to individuals (negatively) affected
by opaque models. An AR recommendations must also be actionable, i.e.,
possible to attain by the recipient. A plausible CE for a rejected
20-year-old loan applicant, for example, might reveal that their
application would have been accepted, if only they had been 20 years
older. Ignoring all other features, this would comply with the
definition of plausibility if 40-year-old individuals were in fact more
credit-worthy on average than young adults. But to the applicant, this
CE does not provide actionable recourse, as age is not a (directly)
mutable feature. Counterfactual training aims to improve model
explainability by aligning models with counterfactuals that meet both
desiderata: plausibility and actionability. Formally, we define
explainability as follows:

\begin{definition}[Model Explainability]
\label{def-explainability}
Let $\mathbf{M}_\theta: \mathcal{X} \mapsto \mathcal{Y}$ denote a supervised classification model that maps from the $D$-dimensional input space $\mathcal{X}$ to representations $\phi(\mathbf{x};\theta)$ and finally to the $K$-dimensional output space $\mathcal{Y}$. Assume that for any given input-output pair $\{\mathbf{x},\mathbf{y}\}_i$ there exists a counterfactual $\mathbf{x}^{\prime} = \mathbf{x} + \Delta: \mathbf{M}_\theta(\mathbf{x}^{\prime}) = \mathbf{y}^{+} \neq \mathbf{y} = \mathbf{M}_\theta(\mathbf{x})$, where $\arg\max_y{\mathbf{y}^{+}}=y^+$ is the index of the target class.

We say that $\mathbf{M}_\theta$ has an \textbf{explanatory capacity} to the extent that faithfully generated counterfactuals are also plausible and actionable. We define these properties as follows:

\begin{enumerate}
    \item (Faithfulness) $\int^{A} p_\theta(\mathbf{x}^\prime|\mathbf{y}^{+})d\mathbf{x} \rightarrow 1$, where $A$ is some arbitrarily small region around $\mathbf{x}^{\prime}$.
    \item (Plausibility) $\int^{A} p(\mathbf{x}^\prime|\mathbf{y}^{+})d\mathbf{x} \rightarrow 1$; $A$ as specified above.
    \item (Actionability) Perturbations $\Delta$ are subject to some actionability constraints.
\end{enumerate}
and $p_\theta(\mathbf{x}|\mathbf{y}^{+})$ denotes the conditional posterior distribution over inputs. For simplicity, we refer to a model with high explanatory capacity as \textbf{explainable} in this manuscript. 
\end{definition}

\noindent The characterization of faithfulness and plausibility in Def.
\ref{def-explainability} is the same as in \citet{altmeyer2024faithful},
with adapted notation. Intuitively, plausible counterfactuals are
consistent with the data and faithful counterfactuals are consistent
with what the model has learned about the input data. Actionability
constraints in Def. \ref{def-explainability} vary and depend on the
context in which \(\mathbf{M}_\theta\) is deployed. In this work, we
choose to only consider domain and mutability constraints for individual
features \(x_d\) for \(d=1,...,D\). We also limit ourselves to
classification tasks for reasons discussed in
Section~\ref{sec-discussion}.

\subsection{Proposed Objective}\label{proposed-objective}

Let \(\mathbf{x}_t^\prime\) for \(t=0,...,T\) denote a counterfactual
generated through gradient descent over \(T\) iterations as originally
proposed by \citet{wachter2017counterfactual}. In broad terms, searching
for CEs using gradient descent entails optimizing some form of an
objective that balances (1) the classification loss
\(\text{yloss}(\mathbf{M}_\theta(\mathbf{x}),\mathbf{y})\), and (2) one
or more penalty terms \(\lambda_{i}\text{cost}_{i}(\cdot)\). The exact
specification of these penalties induces various properties in the
counterfactual outcomes, and tends to be the key feature that
distinguishes various gradient-based ``generators'' or ``explainers'' in
the literature \citep{altmeyer2023explaining}, including all generators
used in our experiments. We refer the reader to the supplementary
appendix for details.

CT adopts gradient-based CE search during training to generate
on-the-fly model explanations \(\mathbf{x}^\prime\) for training
samples. We use the term \emph{nascent} to denote counterfactuals
\(\mathbf{x}_{t\leq T}^\prime\) that are not yet valid where \(t\)
indicates the last iteration before the label is flipped. We store and
use these interim counterfactuals as adversarial examples. Conversely,
we consider counterfactuals \(\mathbf{x}_T^\prime\) as \emph{mature}
explanations if they have either exhausted all \(T\) iterations or
converged in terms reaching a pre-specified threshold, \(\tau\), for the
predicted probability of the target class:
\(\mathcal{S}(\mathbf{M}_\theta(\mathbf{x}^\prime))[y^+] \geq \tau\),
where \(\mathcal{S}\) is the softmax function.

Formally, we propose the following counterfactual training objective to
train explainable (as in Def. \ref{def-explainability}) models:
\begin{equation}\phantomsection\label{eq-obj}{
\begin{aligned}
&\min_\theta \text{yloss}(\mathbf{M}_\theta(\mathbf{x}),\mathbf{y}) + \lambda_{\text{div}} \text{div}(\mathbf{x}^+,\mathbf{x}_T^\prime,y;\theta) \\+ &\lambda_{\text{adv}} \text{advloss}(\mathbf{M}_\theta(\mathbf{x}_{t\leq T}^\prime),\mathbf{y}) + \lambda_{\text{reg}}\text{ridge}(\mathbf{x}^+,\mathbf{x}_T^\prime,y;\theta)
\end{aligned}
}\end{equation} where \(\text{yloss}(\cdot)\) is any classification loss
that induces discriminative performance (e.g., cross-entropy). The
second and third terms are explained in detail below. For now, they can
be summarized as inducing explainability directly and indirectly by
penalizing the contrastive divergence, \(\text{div}(\cdot)\), between
mature counterfactuals \(\mathbf{x}_T^\prime\) and observed samples
\(\mathbf{x}^+\in\mathcal{X}^+=\{\mathbf{x}:y=y^+\}\) in the target
class \(y^+\), and the adversarial loss, \(\text{advloss}(.)\), wrt.
nascent counterfactuals \(\mathbf{x}_{t\leq T}^\prime\). Finally,
\(\text{ridge}(\cdot)\) denotes a Ridge penalty (\(\ell_2\)-norm) that
regularizes the magnitude of the energy terms involved in
\(\text{div}(\cdot)\) \citep{du2019implicit}. The trade-off between the
components are governed through \(\lambda_{\text{div}}\),
\(\lambda_{\text{adv}}\) and \(\lambda_{\text{reg}}\).

\subsection{Directly Inducing Explainability with Contrastive
Divergence}\label{directly-inducing-explainability-with-contrastive-divergence}

\citet{grathwohl2020your} observe that any classifier can be
re-interpreted as a joint energy-based model that learns to discriminate
output classes conditional on the observed (training) samples from
\(p(\mathbf{x})\) and the generated samples from
\(p_\theta(\mathbf{x})\). The authors show that JEMs can be trained to
perform well at both tasks by directly maximizing the joint
log-likelihood:
\(\log p_\theta(\mathbf{x},\mathbf{y})=\log p_\theta(\mathbf{y}|\mathbf{x}) + \log p_\theta(\mathbf{x})\),
where the first term can be optimized using cross-entropy as in
Equation~\ref{eq-obj}. To optimize \(\log p_\theta(\mathbf{x})\), they
minimize the contrastive divergence between the observed samples from
\(p(\mathbf{x})\) and the generated samples from
\(p_\theta(\mathbf{x})\).

A key empirical finding of \citet{altmeyer2024faithful} was that JEMs
perform well on the plausibility objective in Def.
\ref{def-explainability}. This follows directly if we consider samples
drawn from \(p_\theta(\mathbf{x})\) as counterfactuals --- the JEM
objective effectively minimizes the divergence between the conditional
posterior and \(p(\mathbf{x}|\mathbf{y}^{+})\). To generate samples,
\citet{grathwohl2020your} use Stochastic Gradient Langevin Dynamics
(SGLD) with an uninformative prior for initialization but we depart from
their methodology. Instead we propose to leverage counterfactual
explainers to generate counterfactuals of observed training samples.
Specifically, we have: \begin{equation}\phantomsection\label{eq-div}{
\text{div}(\mathbf{x}^+,\mathbf{x}_T^\prime,y;\theta) = \mathcal{E}_\theta(\mathbf{x}^+,y) - \mathcal{E}_\theta(\mathbf{x}_T^\prime,y)
}\end{equation} where \(\mathcal{E}_\theta(\cdot)\) denotes the energy
function defined as
\(\mathcal{E}_\theta(\mathbf{x},y)=-\mathbf{M}_\theta(\mathbf{x})[y^+]\),
with \(y^+\) denoting the index of the randomly drawn target class,
\(y^+ \sim p(y)\). Conditional on the target class \(y^+\),
\(\mathbf{x}_T^\prime\) denotes a mature counterfactual for a randomly
sampled factual from a non-target class generated with a gradient-based
CE generator for up to \(T\) iterations. Mature counterfactuals are ones
that have either reached convergence wrt. the decision threshold
\(\tau\) or exhausted \(T\).

Intuitively, the gradient of Equation~\ref{eq-div} decreases the energy
of observed training samples (positive samples) while increasing the
energy of counterfactuals (negative samples) \citep{du2019implicit}. As
the counterfactuals get more plausible (Def. \ref{def-explainability})
during training, these opposing effects gradually balance each other out
\citep{lippe2024uvadlc}.

The departure from SGLD allows us to tap into the vast repertoire of
explainers that have been proposed in the literature to meet different
desiderata. For example, many methods support domain and mutability
constraints. In principle, any existing approach for generating CEs is
viable, so long as it does not violate the faithfulness condition. Like
JEMs \citep{murphy2022probabilistic}, Counterfactual Training can be
considered a form of contrastive representation learning.

\subsection{Indirectly Inducing Explainability with Adversarial
Robustness}\label{indirectly-inducing-explainability-with-adversarial-robustness}

Based on our analysis in Section~\ref{sec-lit}, counterfactuals
\(\mathbf{x}^\prime\) can be repurposed as additional training samples
\citep{balashankar2023improving, luu2023counterfactual} or adversarial
examples \citep{freiesleben2022intriguing, pawelczyk2022exploring}. This
leaves some flexibility with regards to the choice for the
\(\text{advloss}(\cdot)\) term in Equation~\ref{eq-obj}. An intuitive
functional form, but likely not the only sensible choice, is inspired by
adversarial training: \begin{equation}\phantomsection\label{eq-adv}{
\begin{aligned}
\text{advloss}(\mathbf{M}_\theta(\mathbf{x}_{t\leq T}^\prime),\mathbf{y};\varepsilon)&=\text{yloss}(\mathbf{M}_\theta(\mathbf{x}_{t_\varepsilon}^\prime),\mathbf{y}) \\
t_\varepsilon &= \max_t \{t: ||\Delta_t||_\infty < \varepsilon\}
\end{aligned}
}\end{equation} Under this choice, we consider nascent counterfactuals
\(\mathbf{x}_{t\leq T}^\prime\) as AEs as long as the magnitude of the
perturbation to any single feature is at most \(\varepsilon\). This is
closely aligned with \citet{szegedy2013intriguing} who define an
adversarial attack as an ``imperceptible non-random perturbation''.
Thus, we work with a different distinction between CE and AE than
\citet{freiesleben2022intriguing} who considers misclassification as the
distinguishing feature of adversarial examples. One of the key
observations of this work is that we can leverage CEs during training
and get AEs essentially for free to reap the aforementioned benefits of
adversarial training.

\subsection{Encoding Actionability Constraints}\label{sec-constraints}

Many existing counterfactual explainers support domain and mutability
constraints out-of-the-box. In fact, both types of constraints can be
implemented for any explainer that relies on gradient descent in the
feature space for optimization \citep{altmeyer2023explaining}. In this
context, domain constraints can be imposed by simply projecting
counterfactuals back to the specified domain, if the previous gradient
step resulted in updated feature values that were out-of-domain.
Similarly, mutability constraints can be enforced by setting partial
derivatives to zero to ensure that features are only perturbed in the
allowed direction, if at all.

Since actionability constraints are binding at test time, we should also
impose them when generating \(\mathbf{x}^\prime\) during each training
iteration to inform model representations. Through their effect on
\(\mathbf{x}^\prime\), both types of constraints influence model
outcomes via Equation~\ref{eq-div}. Here it is crucial that we avoid
penalizing implausibility that arises due to mutability constraints. For
any mutability-constrained feature \(d\) this can be achieved by
enforcing \(\mathbf{x}^+[d] - \mathbf{x}^\prime[d]:=0\) whenever
perturbing \(\mathbf{x}^\prime[d]\) in the direction of
\(\mathbf{x}^+[d]\) would violate mutability constraints. Specifically,
we set \(\mathbf{x}^+[d] := \mathbf{x}^\prime[d]\) if:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Feature \(d\) is strictly immutable in practice.
\item
  \(\mathbf{x}^+[d]>\mathbf{x}^\prime[d]\), but \(d\) can only be
  decreased in practice.
\item
  \(\mathbf{x}^+[d]<\mathbf{x}^\prime[d]\), but \(d\) can only be
  increased in practice.
\end{enumerate}

\noindent From a Bayesian perspective, setting
\(\mathbf{x}^+[d] := \mathbf{x}^\prime[d]\) can be understood as
assuming a point mass prior for \(p(\mathbf{x}^+)\) wrt. feature \(d\).
Intuitively, we think of this as ignoring implausibility costs of
immutable features, which effectively forces the model to instead seek
plausibility through the remaining features. This can be expected to
result in lower overall sensitivity to immutable features, which we
investigate empirically in Section~\ref{sec-experiments}. Under certain
conditions, this result holds theoretically; for the proof, see the
supplementary appendix:

\begin{proposition}[Protecting Immutable Features]
\label{prp-mtblty}
Let $f_\theta(\mathbf{x})=\mathcal{S}(\mathbf{M}_\theta(\mathbf{x}))=\mathcal{S}(\Theta\mathbf{x})$ denote a linear classifier with softmax activation $\mathcal{S}$ where $y\in\{1,...,K\}=\mathcal{K}$ and $\mathbf{x} \in \mathbb{R}^D$. Assume multivariate Gaussian class densities with common diagonal covariance matrix $\Sigma_k=\Sigma$ for all $k \in \mathcal{K}$, then protecting an immutable feature from the contrastive divergence penalty will result in lower classifier sensitivity to that feature relative to the remaining features, provided that at least one of those is discriminative and mutable.
\end{proposition}

\section{Experiments}\label{sec-experiments}

We seek to answer the following three research questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  To what extent does the CT objective in Equation 1 induce models to
  learn plausible explanations?
\item
  To what extent does CT lead to more favorable AR outcomes in the
  presence of actionability constraints?
\item
  What are the effects of hyperparameter selection on CT?
\end{enumerate}

\subsection{Experimental Setup}\label{experimental-setup}

Our focus is the improvement in explainability (Def.
\ref{def-explainability}). Thus, we primarily look at the plausibility
and cost of faithfully generated counterfactuals at test time. Other
metrics, such as validity and redundancy, are reported in the
supplementary appendix. To measure the cost, we follow the standard
proxy of distances (\(\ell_1\)-norm) between factuals and
counterfactuals. For plausibility, we assess how similar CEs are to the
observed samples in the target domain,
\(\mathbf{X}^+\subset\mathcal{X}^+\). We rely on the metric used by
\citet{altmeyer2024faithful},
\begin{equation}\phantomsection\label{eq-impl-dist}{
\text{IP}(\mathbf{x}^\prime,\mathbf{X}^+) = \frac{1}{\lvert\mathbf{X}^+\rvert}\sum_{\mathbf{x} \in \mathbf{X}^+} \text{dist}(\mathbf{x}^{\prime},\mathbf{x})
}\end{equation} and introduce a novel divergence metric,
\begin{equation}\phantomsection\label{eq-impl-div}{
\text{IP}^*(\mathbf{X}^\prime,\mathbf{X}^+) = \text{MMD}(\mathbf{X}^\prime,\mathbf{X}^+)
}\end{equation} where \(\mathbf{X}^\prime\) denotes a collection of
counterfactuals and \(\text{MMD}(\cdot)\) is the unbiased estimate of
the squared population maximum mean discrepancy, proposed by
\citet{gretton2012kernel}. The metric in Equation~\ref{eq-impl-div} is
equal to zero if and only if the two distributions are exactly the same,
\(\mathbf{X}^\prime=\mathbf{X}^+\).

For predictive performance, we use standard metrics, such as robust
accuracy estimated on adversarially perturbed data using FGSM
\citep{goodfellow2014explaining}.

We run experiments with three gradient-based generators: \emph{Generic}
of \citet{wachter2017counterfactual} as a simple baseline approach,
\emph{REVISE} \citep{joshi2019realistic} that aims to generate plausible
counterfactuals using a surrogate Variational Autoencoder (VAE), and
\emph{ECCo}---the generator of \citet{altmeyer2024faithful} without the
conformal prediction component---as a method that directly targets both
faithfulness and plausibility of the counterfactuals.

We make use of nine classification datasets common in the CE/AR
literature. Four of them are synthetic with two classes and different
characteristics: linearly separable clusters (\emph{LS}), overlapping
clusters (\emph{OL}), concentric circles (\emph{Circ}), and interlocking
moons (\emph{Moon}). They are generated using the library of
\citet{altmeyer2023explaining} and we present them in the supplementary
appendix. Next, we have four real-world binary tabular datasets:
\emph{Adult} (Census data) of \citet{becker1996adult2}, California
housing (\emph{CH}) of \citet{pace1997sparse}, Default of Credit Card
Clients (\emph{Cred}) of \citet{yeh2016default}, and Give Me Some Credit
(\emph{GMSC}) from \citet{kaggle2011give2}. Finally, for the convenience
of illustration, we use the 10-class \emph{MNIST}
\citep{lecun1998mnist}.

To assess CT, we investigate the improvements in performance metrics
when using it on top of a weak baseline (BL): a multilayer perceptron
(\emph{MLP}). This is the best way to get a clear picture of the
effectiveness of CT, and it is consistent with evaluation practices in
the related literature
\citep{goodfellow2014explaining, ross2021learning, teney2020learning}.

\subsection{Experimental Results}\label{experimental-results}

Our main quantitative results for \emph{MLP} models are summarised in
Table~\ref{tbl-main}, which presents average outcomes along with
bootstrapped two standard errors. The following example motivates CT and
illustrates how to read Table~\ref{tbl-main}.

\subsubsection{Plausibility.}\label{sec-plaus}

We find that CT generally leads to substantial and statistically
significant improvements in plausibility: average reductions in
\(\text{IP}\) range from around 7\% for \emph{MNIST} to almost 60\% for
\emph{Circ}; for the real-world tabular datasets they are around 12\%
for both \emph{CH} and \emph{Cred} and almost 25\% for \emph{GMSC}; for
\emph{Adult} and \emph{OL} we find no significant impact of CT on
\(\text{IP}\). Reductions in \(\text{IP}^*\) are even more substantial
and generally statistically significant, although the average degree of
uncertainty is higher than for \(\text{IP}\): average reductions range
from around 20\% (\emph{Moons}) to almost 90\% (\emph{Circ}). The only
negative findings for \emph{OL} and \emph{MNIST} are statistically
insignificant and, for MNIST, do not align with qualitative findings,
which are much more plausible for CT (Figure~\ref{fig-mnist}).

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../../paper/figures/mnist_body.png}}

}

\caption{\label{fig-mnist}Sample explanations for \emph{MNIST} for BL
(top) and CT (bottom). First column is a random factual 0 (blue).
Columns 2 to 5 are corresponding \emph{ECCo} counterfactuals in target
classes 1 to 4. Columns 6 to 10 show integrated gradients averaged over
all test images in classes 5 to 9.}

\end{figure}%

\subsubsection{Actionability.}\label{sec-act}

We also find that CT can reduce sensitivity to immutable, protected
features and thus lead to less costly counterfactual outcomes as shown
in Figure~\ref{fig-poc}. In column three of Table~\ref{tbl-main}, we
impose mutability constraints on selected features and compute the
reduction in average costs of CEs associated with CT compared to the
baseline: for synthetic datasets, we always protect the first feature;
for all real-world tabular datasets we could identify and protect an
\emph{age} variable; for \emph{MNIST}, we protect the five upper and
lower pixel rows of the full image. Reductions in costs are
overwhelmingly positive and significant of up to nearly 60\% for
\emph{GMSC}. While the estimated cost reductions for \emph{Adult} and
\emph{MNIST} are not significant, Figure~\ref{fig-mnist} (columns 6-10)
demonstrates that CT does have the expected effect: sensitivity to
protected features as per integrated gradients is drastically reduced;
details of this experiment are reported in the supplementary appendix.
In the case of \emph{Cred}, average costs increase, likely because any
potential benefits from protecting the \emph{age} are outweighed by the
increase in costs required for greater plausibility.

\subsubsection{Predictive Performance.}\label{sec-pred}

Test accuracy for CT is virtually identical to the baseline for
\emph{Adult}, \emph{Circ}, \emph{LS}, \emph{Moon}, and \emph{OL}, and
even slightly improved for \emph{Cred}. Exceptions to this general
pattern are \emph{MNIST}, \emph{CH}, and \emph{GMSC}, for which we
observe a reduction in test accuracy of 2, 5, and 15 percentage points
respectively. When looking at robust test accuracies (Acc.\(^*\)) for
these datasets in particular, we find that CT strongly outperforms the
baseline. In fact, we observe that CT improves adversarial robustness on
all datasets.

\begin{table*}

\caption{\label{tbl-main}Key performance metrics and bootstrapped
standard errors (omitted if essentially zero) for all datasets.
\textbf{Plausibility} (columns 1-2): percentage reduction in
implausibility for \(\text{IP}\) and \(\text{IP}^*\), respectively;
\textbf{Actionability} (column 3): percentage reduction in costs with
protected features. \textbf{Accuracy} (columns 4-7): test accuracies and
robust accuracies (\(\text{Acc}^*\)) for CT and the baseline (BL).
Counterfactual outcomes (col.~1-3) are aggregated across bootstrap
samples and varying degrees of the energy penalty
\(\lambda_{\text{egy}}\) used for \emph{ECCo} at test time. Asterisks
(\(^*\)) in columns 1-3 indicate that bootstrapped 99\%-confidence
interval of differences in mean outcomes does \emph{not} include zero.
Standard errors for accuracy (col.~4-7) are bootstrapped from the test
set.}

\centering{

\input{tables/main.tex}

}

\end{table*}%

\subsubsection{Hyperparameter settings.}\label{sec-hyperparameters}

We test the impact of three types of hyperparameters. Here we focus on
the highlights; full results are available in the supplementary
appendix.\\
\indent First, we note that CT is highly sensitive to the choice of a CE
generator and its hyperparameters but (1) there are manageable patterns,
and (2) we can usually identify settings that improve either
plausibility or cost, and often both of them at the same time. For
example, \emph{REVISE} tends to perform the worst, most likely because
it uses a surrogate VAE to generate counterfactuals which impedes
faithfulness \citep{altmeyer2024faithful}. Increasing \(T\), the maximum
number of steps, generally yields better outcomes because more CEs can
mature in each training epoch. The impact of \(\tau\), the required
decision threshold is more difficult to predict. On ``harder'' datasets
it may be difficult to satisfy high \(\tau\) for any given sample (i.e.,
also factuals) and so increasing this threshold does not seem to
correlate with better outcomes. In fact, \(\tau=0.5\) generally leads to
optimal results as it is associated with high proportions of mature
counterfactuals.\\
\indent Second, the strength of the energy regularization,
\(\lambda_{\text{reg}}\) is highly impactful and leads to poor
performance in terms of decreased plausibility and increased costs if
insufficiently high. The sensitivity with respect to
\(\lambda_{\text{div}}\) and \(\lambda_{\text{adv}}\) is much less
evident. While high values of \(\lambda_{\text{reg}}\) may increase the
variability in outcomes when combined with high values of
\(\lambda_{\text{div}}\) or \(\lambda_{\text{adv}}\), this effect is not
very pronounced.\\
\indent Third, the effectiveness and stability of CT is positively
associated with the number of counterfactuals generated during each
training epoch. A higher number of training epochs is also beneficial.
Interestingly, we observed desired improvements when CT was combined
with conventional training and applied only for the final 50\% of epochs
of the complete training process. Put differently, CT can improve the
explainability of models in a fine-tuning manner.

\section{Conclusions}\label{sec-discussion}

As our results indicate, counterfactual training produces models that
are more explainable. Nonetheless, these advantages come at the cost of
two important limitations.

\emph{Interventions on features have implications for fairness.} We
provide a tool that allows practitioners to modify the sensitivity of a
model with respect to certain features. Model owners can use our
solution to support the fair and equitable treatment of decision
subjects, but they could also misuse it by enforcing explanations based
on features that are more difficult to modify by some (group of)
individuals. When used irresponsibly, CT could result in an unfairly
assigned burden of recourse \citep{sharma2020certifai}, threatening the
equality of opportunity \citep{bell2024fairness}. Additionally, CT
requires mutability constraints for the features considered by the
model. Even if all immutable features are protected, there may exist
proxies that are mutable, and hence should not be protected, but
preserve sufficient information about the principals to hinder these
protections. Deciding on actionability is still a major open challenge
in the AR literature \citep{venkatasubramanian2020philosophical}
impacting the capacity of CT to fulfill its intended goal.

\emph{CT increases the training times.} Like adversarial training, CT is
more resource-intensive than conventional regimes. Higher numbers of CEs
improve the quality of learned representations but they also increase
the number of computations. As our codebase is not performance
optimized, grids of 270 settings for the largest datasets in our
experiments took up to four hours using 34 2GB CPUs (see supplementary
appendix). Other than optimization, three factors mitigate this effect:
(1) CT yields itself to parallel execution; (2) it amortizes the cost of
CEs for the training samples; and (3) it can be used to fine-tune
conventionally-trained models.

We also highlight three important directions for future research.
Firstly, it is an interesting challenge to extend CT beyond
classification settings. Our formulation relies on the distinction
between non-target class(es) \(y^{-}\) and target class(es) \(y^{+}\) to
generate counterfactuals through Equation~\ref{eq-obj}. While \(y^{-}\)
and \(y^{+}\) can be arbitrarily defined, CT requires the output space
\(\mathcal{Y}\) to be discrete. Thus, it does not apply to ML tasks
where the change in outcome cannot be readily discretized. Focus on
classification is a common choice in research on CEs and AR. Other
settings have attracted some interest, e.g., regression
\citep{spooner2021counterfactual}, but there is little consensus how to
robustly extend the notion of CEs.\\
Secondly, CT is susceptible to training instabilities. This problem has
been recognized for JEMs \citep{grathwohl2020your} and even though we
depart from the SGLD sampling, we still encounter variability in
outcomes. CT is exposed to two potential sources of instabilities: (1)
the energy-based contrastive divergence term, \(\text{div}(\cdot)\), in
Equation~\ref{eq-div}, and (2) the underlying explainers. We find
several promising ways to mitigate this problem: regularizing energy
(\(\lambda_{\text{reg}}\)), generating sufficiently many counterfactuals
during each epoch, and including only mature counterfactuals in
\(\text{div}(\cdot)\).\\
Finally, we believe that it is possible to considerably improve
hyperparameter selection procedures, and thus performance. We have
relied exclusively on grid searches, but future research could benefit
from more sophisticated approaches.

To conclude, state-of-the-art machine learning models are prone to
learning complex representations that cannot be interpreted by humans.
Existing explainability solutions cannot guarantee that explanations
agree with the model's learned representation of data. As a step towards
addressing this challenge, we introduced counterfactual training, a
novel training regime that incentivizes highly-explainable models. Our
approach leads to explanations that are both plausible (compliant with
the underlying data-generating process) and actionable (compliant with
user-specified mutability constraints), and thus meaningful to their
recipients. Through extensive experiments, we demonstrated that CT
satisfies its objective while promoting robustness and preserving the
predictive performance of models. It can also be used to fine-tune
conventionally-trained models and achieve similar gains. Lastly, our
work highlights the value of simultaneously improving models and their
explanations.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{../quarto_aaai/bibliography.bib}
\end{document}