### 2025-07-18 paper/sections/experiments.qmd
final adjustments to text

# Experiments {#sec-experiments}

We seek to answer the following three research questions:

1. To what extent does the CT objective in Equation 1 induce models to learn plausible explanations?
2. To what extent does CT lead to more favorable AR outcomes in the presence of actionability constraints?
3. What are the effects of hyperparameter selection on CT?

## Experimental Setup

Our focus is the improvement in explainability (Def. \ref{def-explainability}). Thus, we primarily look at the plausibility and cost of faithfully generated counterfactuals at test time. Other metrics, such as validity and redundancy, are reported in the supplementary appendix. To measure the cost, we follow the standard proxy of distances ($\ell_1$-norm) between factuals and counterfactuals. For plausibility, we assess how similar CEs are to the observed samples in the target domain, $\mathbf{X}^+\subset\mathcal{X}^+$. We rely on the metric used by @altmeyer2024faithful,
$$
\text{IP}(\mathbf{x}^\prime,\mathbf{X}^+) = \frac{1}{\lvert\mathbf{X}^+\rvert}\sum_{\mathbf{x} \in \mathbf{X}^+} \text{dist}(\mathbf{x}^{\prime},\mathbf{x})
$$ {#eq-impl-dist}
and introduce a novel divergence-based adaption,
$$
\text{IP}^*(\mathbf{X}^\prime,\mathbf{X}^+) = \text{MMD}(\mathbf{X}^\prime,\mathbf{X}^+)
$$ {#eq-impl-div}
where $\mathbf{X}^\prime$ denotes a collection of counterfactuals and $\text{MMD}(\cdot)$ is the unbiased estimate of the squared population maximum mean discrepancy, proposed by @gretton2012kernel. The metric in @eq-impl-div is equal to zero if and only if the two distributions are exactly the same, $\mathbf{X}^\prime=\mathbf{X}^+$.

To assess outcomes with respect to actionability for non-linear models, we look at the average costs of valid counterfactuals in terms of their distances from factual starting points. While this an imperfect proxy of sensitivity, we hypothesize that CT can reduce these costs by teaching models to seek plausibility with respect to mutable features, much like we observe in @fig-poc (panel (d) vs panel (c)). We supplement this analysis with qualitative findings for integrated gradients [@sundararajan2017ig]. Finally, for predictive performance, we use standard metrics, such as robust accuracy estimated on adversarially perturbed data using FGSM [@goodfellow2014explaining].

We run experiments with three gradient-based generators: *Generic* of @wachter2017counterfactual as a simple baseline approach, *REVISE* [@joshi2019realistic] that aims to generate plausible counterfactuals using a surrogate Variational Autoencoder (VAE), and *ECCo*---the generator of @altmeyer2024faithful without the conformal prediction component---as a method that directly targets both faithfulness and plausibility of the counterfactuals.

We make use of nine classification datasets common in the CE/AR literature. Four of them are synthetic with two classes and different characteristics: linearly separable clusters (*LS*), overlapping clusters (*OL*), concentric circles (*Circ*), and interlocking moons (*Moon*). They are generated using the library of @altmeyer2023explaining and we present them in the supplementary appendix. Next, we have four real-world binary tabular datasets: *Adult* (Census data) of @becker1996adult2, California housing (*CH*) of @pace1997sparse, Default of Credit Card Clients (*Cred*) of @yeh2016default, and Give Me Some Credit (*GMSC*) from @kaggle2011give2. Finally, for the convenience of illustration, we use the 10-class *MNIST* [@lecun1998mnist].

To assess CT, we investigate the improvements in performance metrics when using it on top of a weak baseline (BL): a multilayer perceptron (*MLP*). This is the best way to get a clear picture of the effectiveness of CT, and it is consistent with evaluation practices in the related literature [@goodfellow2014explaining;@ross2021learning;@teney2020learning].

## Experimental Results

Our main results for plausibility and actionability for *MLP* models are summarised in @tbl-main, which presents counterfactual outcomes grouped by dataset along with standard errors averaged across bootstrap samples.

The first two columns ($\text{IP}$ and $\text{IP}^*$) show the percentage reduction in implausibility for our two metrics when using CT on top of the weak baseline. Focusing on the first row for *LS* for a moment, the observed positive values indicate that faithful counterfactuals are around 30-55% more plausible for models trained with CT, in line with our observations in @fig-poc (panel (b) vs. (a)). Asterisks indicate that reductions are statistically significant: bootstrapped 99%-confidence intervals of differences in mean outcomes do *not* include zero.

The third column shows results for the scenario when mutability constraints are imposed on selected features. Once again we are comparing CT to the baseline, so reductions in the positive direction imply that valid counterfactuals are cheaper when using CT with feature protection. Relating this back to @fig-poc, the third column represents the reduction in distances travelled by counterfactuals in panel (d) compared to (c). In the following, we summarize the remaining results.

::: {#tbl-main}

```{=latex}
\begin{table}[h]
\small
\centering
\input{tables/main.tex}
\end{table}
```

Key evaluation metrics for valid counterfactual along with bootstrapped standard errors for all datasets. **Plausibility** (columns 1-2): percentage reduction in implausibility for $\text{IP}$ and $\text{IP}^*$, respectively; **Actionability** (column 3): percentage reduction in costs with protected features. Outcomes are aggregated across bootstrap samples and varying degrees of the energy penalty $\lambda_{\text{egy}}$ used for *ECCo* at test time. Asterisks ($^*$) indicate that bootstrapped 99%-confidence interval of differences in mean outcomes does *not* include zero.

:::

### Plausibility. {#sec-plaus}

We find that CT generally leads to substantial and statistically significant improvements in plausibility: average reductions in $\text{IP}$ range from around 7% for *MNIST* to almost 60% for *Circ*; for the real-world tabular datasets they are around 12% for both *CH* and *Cred* and almost 25% for *GMSC*; for *Adult* and *OL* we find no significant impact of CT on $\text{IP}$. Reductions in $\text{IP}^*$ are even more substantial and generally statistically significant, although the average degree of uncertainty is higher than for $\text{IP}$: average reductions range from around 20% (*Moons*) to almost 90% (*Circ*). The only negative findings for *OL* and *MNIST* are statistically insignificant and, for MNIST, do not align with qualitative findings, which are much more plausible for CT (@fig-mnist).

![Sample explanations for *MNIST* for BL (top) and CT (bottom). First column is a random factual 0 (blue). Columns 2 to 5 are corresponding *ECCo* counterfactuals in target classes 1 to 4. Columns 6 to 10 show integrated gradients averaged over all test images in classes 5 to 9.](/paper/figures/mnist_body.png){#fig-mnist}

### Actionability. {#sec-act}

We also find that CT can improve actionability in the presence of immutable features (column 3 in @tbl-main): for synthetic datasets, we always protect the first feature; for all real-world tabular datasets we could identify and protect an *age* variable; for *MNIST*, we protect the five upper and lower pixel rows of the full image. Reductions in costs are overwhelmingly positive and significant of up to nearly 60% for *GMSC*. While the estimated cost reductions for *Adult* and *MNIST* are not significant, @fig-mnist (columns 6-10) demonstrates that in the latter case, CT does have the expected effect: sensitivity to protected features as per integrated gradients is drastically reduced; details of this experiment are reported in the supplementary appendix. In the case of *Cred*, average costs increase, likely because any potential benefits from protecting the *age* are outweighed by the increase in costs required for greater plausibility.

### Predictive Performance. {#sec-pred}

Test accuracies on adversarially perturbed data are shown in @fig-acc. The perturbations size, $\varepsilon \in [0,0.1]$, increases along the horizontal axis and includes zero, corresponding to standard test accuracy for non-perturbed data. We observe that for all synthetic datasets, predictive performance of CT is virtually identical to the baseline and unaffected by adversarial perturbations. For all real-world datasets, we find that CT substantially improves robustness: while in some cases baseline accuracies drop to essentially zero for large enough perturbation sizes, accuracies for CT are remarkably robust to gradient-based adversarial attacks.

![Test accuracies on adversarially perturbed data with varying perturbation sizes for all non-synthetic datasets.](/paper/figures/acc.png){#fig-acc fig-env="figure*"}

### Hyperparameter settings. {#sec-hyperparameters}
We test the impact of three types of hyperparameters. Here we focus on the highlights; full results are available in the supplementary appendix.  
`\indent`{=latex} First, we note that CT is highly sensitive to the choice of a CE generator and its hyperparameters but (1) there are manageable patterns, and (2) we can usually identify settings that improve either plausibility or cost, and often both of them at the same time. For example, *REVISE* tends to perform the worst, most likely because it uses a surrogate VAE to generate counterfactuals which impedes faithfulness [@altmeyer2024faithful]. Increasing $T$, the maximum number of steps, generally yields better outcomes because more CEs can mature in each training epoch. The impact of $\tau$, the required decision threshold is more difficult to predict. On "harder" datasets it may be difficult to satisfy high $\tau$ for any given sample (i.e., also factuals) and so increasing this threshold does not seem to correlate with better outcomes. In fact, $\tau=0.5$ generally leads to optimal results as it is associated with high proportions of mature counterfactuals.  
`\indent`{=latex} Second, the strength of the energy regularization, $\lambda_{\text{reg}}$ is highly impactful and leads to poor performance in terms of decreased plausibility and increased costs if insufficiently high. The sensitivity with respect to $\lambda_{\text{div}}$ and $\lambda_{\text{adv}}$ is much less evident. While high values of $\lambda_{\text{reg}}$ may increase the variability in outcomes when combined with high values of $\lambda_{\text{div}}$ or $\lambda_{\text{adv}}$, this effect is not very pronounced.   
`\indent`{=latex} Third, the effectiveness and stability of CT is positively associated with the number of counterfactuals generated during each training epoch. A higher number of training epochs is also beneficial. Interestingly, we observed desired improvements when CT was combined with conventional training and applied only for the final 50% of epochs of the complete training process. Put differently, CT can improve the explainability of models in a fine-tuning manner.