### 2025-07-22 paper/sections/discussion.qmd

# Discussion {#sec-discussion}

As our results indicate, counterfactual training produces models that are more explainable. Nonetheless, these advantages come at the cost of two important limitations.

*Interventions on features have implications for fairness.* We provide a method to modify the sensitivity of a model to certain features, which can be misused by enforcing explanations based on features that are more difficult to modify by a (group of) decision subjects. Such abuse could result in an unfairly assigned burden of recourse [@sharma2020certifai], threatening the equality of opportunity [@bell2024fairness]. Also, even if all immutable features are protected, there may exist proxies that are theoretically mutable, but preserve sufficient information about the principals to hinder these protections. Indeed, deciding on the actionability of features remains a major open challenge in the AR literature [@venkatasubramanian2020philosophical].

*Plausibility is costly.* As noted by @altmeyer2024faithful, more plausible counterfactuals are inevitably more costly. CT improves plausibility and robustness, but it can impact average costs and validity when cheap, implausible and adversarial explanations are removed from the solution space.

*CT increases the training times.* Just like contrastive and robust learning, CT is more resource-intensive than conventional regimes. Three factors mitigate this effect: (1) CT yields itself to parallel execution; (2) it amortizes the cost of CEs for the training samples; and (3) it can be used to fine-tune conventionally-trained models.

We also highlight three key directions for future research. Firstly, it is an interesting challenge to extend CT beyond classification settings. Our formulation relies on the distinction between non-target class(es) and target class(es), requiring the output space to be discrete. Thus, it does not apply to ML tasks where the change in outcome cannot be readily discretized. Focus on classification is a common choice in research on CEs and AR; other settings have attracted some interest, e.g., regression [@spooner2021counterfactual], but there is little consensus how to robustly extend the notion of CEs.

Secondly, our analysis covers CE generators with different characteristics, but it is interesting to extend it to more algorithms, including ones that do not rely on computationally costly gradient-based optimization. This should reduce training costs while possibly preserving the benefits of CT.

Finally, we believe that it is possible to considerably improve hyperparameter selection procedures, and thus performance. We have relied exclusively on grid searches, but future work could benefit from more sophisticated approaches.