### 2025-07-22 paper/preprint/appendix.qmd
finished table and rerendered preprint

---
title: Supplementary Appendix
engine: julia
julia: 
  exeflags: ["--project=../experiments/"]
format:
  arxiv-pdf:
    keep-tex: true  
    linenumbers: false
    doublespacing: false
    runninghead: "Counterfactual Training"
    authorcols: true
    include-in-header:
      text: |
        \usepackage[title]{appendix}
        \usepackage{placeins}
        \usepackage{amsthm}
        \theoremstyle{plain}
        \newtheorem{proposition}{Proposition}[section]
        \theoremstyle{definition}
        \newtheorem{definition}{Definition}[section]
        \theoremstyle{definition}
        \newtheorem{example}{Example}[section]
        \theoremstyle{plain}
---

```{=latex}
\begin{appendices}
```

{{< include /paper/sections/appendix.qmd >}}

```{=latex}
\end{appendices}
```