### 2025-07-22 paper/sections/appendix.qmd

{{< include /paper/sections/other/notation.qmd >}}

{{< include /paper/sections/other/technical.qmd >}}

{{< include /paper/sections/other/main_results.qmd >}}

\FloatBarrier

{{< include /paper/sections/other/grids.qmd >}}

\FloatBarrier

{{< include /paper/sections/other/tuning.qmd >}}

\FloatBarrier

{{< include /paper/sections/other/compute.qmd >}}