### 2025-07-23 paper/preprint/appendix.qmd

---
title: Supplementary Appendix
subtitle: "Counterfactual Training: Teaching Models Plausible and Actionable Explanations"
abstract: |
  This is the supplementary appendix to our paper titled *Counterfactual Training: Teaching Models Plausible and Actionable Explanations*. It provides helpful details on mathematical notations and formulas, our proposed training regime, extended empirical findings, hyperparameter tuning and grid searches as well as software and computations.
engine: julia
julia: 
  exeflags: ["--project=../experiments/"]
format:
  arxiv-pdf:
    toc: true
    keep-tex: true  
    linenumbers: false
    doublespacing: false
    runninghead: "Counterfactual Training"
    authorcols: true
    include-in-header:
      text: |
        \usepackage[title]{appendix}
        \usepackage{placeins}
        \usepackage{amsthm}
        \theoremstyle{plain}
        \newtheorem{proposition}{Proposition}[section]
        \theoremstyle{definition}
        \newtheorem{definition}{Definition}[section]
        \theoremstyle{definition}
        \newtheorem{example}{Example}[section]
        \theoremstyle{plain}
---

{{< pagebreak >}}

```{=latex}
\begin{appendices}
```

{{< include /paper/sections/appendix.qmd >}}

```{=latex}
\end{appendices}
```

# References {-}

::: {#refs}
:::