### 2025-07-23 paper/sections/other/main_results.qmd

```{julia}
projectdir = splitpath(pwd()) |>
    ss -> joinpath(ss[1:findall([s == "CounterfactualTraining.jl" for s in ss])[1]]...) 
cd(projectdir)

using CTExperiments
using CTExperiments.CairoMakie
using CTExperiments.CSV
using CTExperiments.DataFrames
using CTExperiments.Plots
using CTExperiments.PrettyTables
using CTExperiments.StatsBase
using Random
using Serialization

using DotEnv
DotEnv.load!()
```

```{julia}
res_dir = joinpath(ENV["FINAL_GRID_RESULTS"],"single")
mtbl_dir = joinpath(ENV["FINAL_GRID_RESULTS"],"mutability")
```

\FloatBarrier

# Details on Main Experiments {#sec-app-main .appendix}

## Final Hyperparameters

As discussed the main paper, CT is sensitive to certain hyperparameter choices. We study the effect of many hyperparameters extensively in @sec-app-grid. For the main results, we tune a small set of key hyperparameters (@sec-app-tune). The final choices for the main results are presented for each data set in @tbl-final-params along with training, test and batch sizes.

::: {#tbl-final-params}

::: {.content-hidden unless-format="pdf"}

```{julia}
#| output: asis

df = final_params(res_dir)
float_cols = ["decision_threshold", "lambda_energy_reg"]
get_table_inputs(df, nothing; backend=Val(:latex)) |>
    inputs -> tabulate_results(inputs; wrap_table=false, table_type=:tabular, formatters=PrettyTables.ft_round(2, findall([x in float_cols for x in names(df)])))
```

:::

Final hyperparameters used for the main results presented in the main paper. Any hyperparameter not shown here is set to its default value (@nte-train-default).

:::

## Final Results {#sec-app-final-results}

Plus/minus two standard deviations of bootstrap estimates.

<!-- Standard deviation of bootstrap is the standard error. -->

```{julia}
#| eval: false

df_mtbl = aggregate_ce_evaluation(
    mtbl_dir; 
    ratio=true, 
    rebase=false, 
    y="distance", 
    agg_further_vars=["run", "lambda_energy_eval"], 
    total_uncertainty=false, 
    drop_models=["linear"]
)
df = final_table(res_dir; tbl_mtbl=df_mtbl)
Serialization.serialize("paper/experiments/output/final_table.jls", df)
```

```{julia}
#| eval: false

df = Serialization.deserialize("paper/experiments/output/final_table.jls")
inputs = get_table_inputs(df, nothing; backend=Val(:latex))
for fname in ["paper/quarto_aaai/tables/main.tex","paper/preprint/tables/main.tex"]
    tabulate_results(
        inputs; 
        table_type=:tabular, 
        save_name=fname, 
        wrap_table=false, 
    )
end
```

### Robust Performance Plots

```{julia}
#| eval: false

plt = CTExperiments.plot_performance(res_dir, eps=range(0.0, 0.1, length=10) |> collect, adversarial=true, byvars=["objective", "eps"], drop_synthetic=false)
plt_out = draw(plt, figure=(size=(1000, 150),), axis=(yticks=[0.0,0.5,1.0], limits=(nothing, (0,1)), xticklabelsvisible=false, xticksvisible=false))
save("paper/figures/acc.png", plt_out; px_per_unit=3)
```

### Confidence Intervals

```{julia}
#| eval: false

lvls = [0.90, 0.95, 0.99]

# Mutability:
y = "distance"
df_mtbl = aggregate_ce_evaluation(
    mtbl_dir; 
    y,
    return_sig_level=false, 
    conf_int=lvls,
    agg_further_vars=["run", "lambda_energy_eval"],
    drop_models=["linear"]
)
df_mtbl.measure .= y

# Distance-based:
y = "plausibility_distance_from_target"
df_ip = aggregate_ce_evaluation(
    res_dir;
    y,
    return_sig_level=false, 
    conf_int=lvls,
    agg_further_vars=["run", "lambda_energy_eval"],
) 
df_ip.measure .= y

# Divergence-based:
y = "mmd"
df_mmd = aggregate_ce_evaluation(
    res_dir;
    y,
    return_sig_level=false, 
    conf_int=lvls,
    agg_further_vars=["run", "lambda_energy_eval"],
)
df_mmd.measure .= y
df = vcat(df_mtbl, df_ip, df_mmd)
Serialization.serialize("paper/experiments/output/ci.jls", df)
```

::: {#tbl-ci}

```{julia}
#| output: asis

df = Serialization.deserialize("paper/experiments/output/ci.jls")
df = select(df, :variable, :measure, :dataset, :full, :vanilla, "99.0%")
df.variable = (x -> LatexCell(replace(x.data, " \$(-%)\$" => ""))).(df.variable)
sort!(df, [:measure, :dataset])
df = select(df, Not(:measure))
df = transform(df, "99.0%" => AsTable) |>
    df -> select(df, Not("99.0%"))
rename!(df, :dataset => :data, :x1 => :LB, :x2 => :UB, :full => :CT, :vanilla => :BL)
inputs = get_table_inputs(df, nothing; backend=Val(:latex))
tabulate_results(
    inputs; 
    table_type=:tabular,
    alignment = [:l, fill(:c, ncol(df)-1)...],
    wrap_table = false,
    sort_by_data = false,
    hlines = :none,
)

```

Mean outcomes for **CT** and **BL** along with bootstrapped confidence intervals (99%) for difference in mean outcomes grouped by dataset and evaluation metric. Column **LB** and **UB** show the lower and upper bound of the intervals, respectively, and computed using the percentile method. The underlying counterfactual evaluations are the same as the ones used to produce the main table in the paper.

:::

### Qualitative Findings for Image Data

```{julia}
#| eval: false

Random.seed!(42)    # change seed for different outcome
overwrite = false   # set to `true` if you want to overwrite the file
fname = joinpath(res_dir,"mlp/mnist/grid_config.toml")
output_dir = mkpath("paper/experiments/output/extra/")
gen = CTExperiments.GeneratorParams(lambda_energy=25.0)     # instead of a high penalty, you can also use e.g. 1.0 and many iterations e.g 1000
conv = "max_iter"
cfg = EvaluationConfig(
    grid_file=fname, 
    counterfactual_params=(conv=conv, generator_params=gen), 
    save_dir=output_dir,
    test_time=true
)
generate_factual_target_pairs(cfg; overwrite, nce=10)
plts = plot_ce(MNIST(), cfg; byvars="objective")
savefig(plts["objective"]["full"]["ecco"], "paper/figures/mnist_mlp.png")
savefig(plts["objective"]["vanilla"]["ecco"], "paper/figures/mnist_mlp_vanilla.png")
```

@fig-mnist shows much more plausible (faithful) counterfactuals for a model with CT than the model with conventional training (@fig-mnist-vanilla).

::: {layout="[10,-2,10]" layout-valign="top"}
![Counterfactual images for *MLP* with counterfactual training. Factual images are shown on the diagonal, with the corresponding counterfactual for each target class (columns) in that same row. The underlying generator, *ECCCo*, aims to generate counterfactuals that are faithful to the model [@altmeyer2024faithful].](/paper/figures/mnist_mlp.png){#fig-mnist}

![The same setup, factuals, model architecture and generator as in @fig-mnist, but the model was trained conventionally.](/paper/figures/mnist_mlp_vanilla.png){#fig-mnist-vanilla}
:::

::: {layout-ncol=2}

### Integrated Gradients

:::{#tbl-ig}

```{julia}
#| output: asis

df = Serialization.deserialize(joinpath(mtbl_dir, "ig.jls")) |>
    df -> transform(df, [:mean, :se] => ByRow((m, s) -> LatexCell("$(round(m, digits=2)) \\pm $(round(s, digits=2))")) => :mean) |>
    df -> select(df, Not(:se)) |>
    df -> DataFrames.unstack(df, :objective, :mean) |>
    df -> rename(df, :full => :CT, :vanilla => :BL) |>
    df -> transform(df, :data => ByRow(d  -> CTExperiments.format_header(d)) => :data)
inputs = get_table_inputs(df, nothing; backend=Val(:latex))
al = [
    "l", 
    "S[table-format=2.2(3.2)]", 
    "S[table-format=2.2(2.2)]"
] 
tabulate_results(
    inputs,
    al
    ; 
    table_type=:tabular,
    alignment = [:l, fill(:c, ncol(df)-1)...],
    wrap_table = false,
)

```

Integrated gradients.

:::

![Class-conditional integrated gradients.](/paper/figures/mnist_ig.png){#fig-mnist-ig}

:::

### Costs and Validity

::: {#tbl-panel layout-ncol=3}

```{julia}
#| eval: false

df = final_table(
    res_dir; 
    ce_var=["distance"], 
    agg_further_vars=[["run", "lambda_energy_eval"]], 
    total_uncertainty=false, 
)
Serialization.serialize("paper/experiments/output/costs.jls", df)
```

::: {#tbl-costs}

```{julia}
#| output: asis
#| eval: true

df = Serialization.deserialize("paper/experiments/output/costs.jls")
inputs = get_table_inputs(df, nothing; backend=Val(:latex))
al = [
    "l", 
    "S[table-format=2.2(1.2)]", 
]
tabulate_results(
    inputs, al; 
    table_type=:tabular,
    alignment = [:l, fill(:c, ncol(df)-1)...],
    wrap_table = false,
)

```

Costs

:::

```{julia}
#| eval: false

# Unconstrained:
df = aggregate_ce_evaluation(
    res_dir; 
    ratio=false, 
    rebase=false, 
    y="validity", 
    agg_further_vars=["run", "lambda_energy_eval"], 
    total_uncertainty=false, 
    drop_models=["linear"]
)
df = select(df, Not(:se, :variable)) |> df -> unstack(df, :objective, :mean)
rename!(df, :dataset => :data, :Full => :CT, :Vanilla => :BL)
Serialization.serialize("paper/experiments/output/validity.jls", df)

# Mutability constrained:
df = aggregate_ce_evaluation(
    mtbl_dir; 
    ratio=false, 
    rebase=false, 
    y="validity", 
    agg_further_vars=["run", "lambda_energy_eval"], 
    total_uncertainty=false, 
    drop_models=["linear"]
)
df = select(df, Not(:se, :variable)) |> df -> unstack(df, :objective, :mean)
rename!(df, :dataset => :data, :Full => :CT, :Vanilla => :BL)
Serialization.serialize("paper/experiments/output/validity_mtbl.jls", df)
```

::: {#tbl-val}

```{julia}
#| output: asis
#| eval: true

df = Serialization.deserialize("paper/experiments/output/validity.jls")
inputs = get_table_inputs(df, nothing; backend=Val(:latex))
tabulate_results(
    inputs; 
    table_type=:tabular,
    alignment = [:l, fill(:c, ncol(df)-1)...],
    wrap_table = false,
)

```

Validity

:::

::: {#tbl-val-mtbl}

```{julia}
#| output: asis
#| eval: true

df = Serialization.deserialize("paper/experiments/output/validity_mtbl.jls")
inputs = get_table_inputs(df, nothing; backend=Val(:latex))
tabulate_results(
    inputs; 
    table_type=:tabular,
    alignment = [:l, fill(:c, ncol(df)-1)...],
    wrap_table = false,
)

```

Validity

:::

Costs and validity.

:::