### 2025-07-23 paper/sections/other/compute.qmd

# Computation Details {.appendix}

<!-- [@DHPC2022] -->
## Hardware {#sec-app-hardware}

We performed our experiments on a high-performance cluster. Details about the cluster will be disclosed upon publication to avoid revealing information that might interfere with the double-blind review process. Since our experiments involve highly parallel tasks and rather small models by today's standard, we have relied on distributed computing across multiple central processing units (CPU). Graphical processing units (GPU) were *not* used.

### Grid Searches

Model training for the largest grid searches with 270 unique parameter combinations was parallelized across 34 CPUs with 2GB memory each. The time to completion varied by dataset: 0h49m (*Moons*), 1h4m (*Linearly Separable*), 1h49m (*Circles*), 3h52m (*Overlapping*). Model evaluations for large grid searches were parallelized across 20 CPUs with 3GB memory each. Evaluations for all data sets took less than one hour (<1h) to complete but were generally more memory-intensive (see @sec-app-software for additional details)

### Tuning

For tuning of selected hyperparameters, we distributed the task of generating counterfactuals during training across 40 CPUs with 2GB memory each for all tabular datasets. Except for the *Adult* dataset, all training runs were completed in less that half an hour (<0h30m). The *Adult* dataset took around 0h35m to complete. Evaluations across 20 CPUs with 3GB memory each generally took less than 0h30m to complete. For *MNIST*, we relied on 100 CPUs with 2GB memory each. For the *MLP*, training of all models could be completed in 1h30m, while the evaluation across 20 CPUs (6GB memory) took 4h12m. For the *CNN*, training of all models took ~8h, with conventionally trained models taking ~0h15m each and model with CT taking ~0h30m-0h45m each.

## Software {#sec-app-software}

All computations were performed in the Julia Programming Language [@bezanson2017julia]. We have developed a package for counterfactual training that leverages and extends the functionality provided by several existing packages, most notably [CounterfactualExplanations.jl](https://github.com/JuliaTrustworthyAI/CounterfactualExplanations.jl) [@altmeyer2023explaining] and the [Flux.jl](https://fluxml.ai/Flux.jl/v0.16/) library for deep learning [@innes2018fashionable;@innes2018flux]. We chose to work with [CounterfactualExplanations.jl](https://github.com/JuliaTrustworthyAI/CounterfactualExplanations.jl) because it currently appears to be the most comprehensive and extensible package for counterfactual explanations. Despite its good interplay with Flux.jl, the package is not, however, optimized to be used in training. This has caused some issues with memory management and bottlenecked performance. All of our code will be provided to reviewers of this work and open-sourced upon publication.

For data-wrangling and presentation-ready tables we relied on [DataFrames.jl](https://dataframes.juliadata.org/v1.7/) [@milan2023dataframes] and [PrettyTables.jl](https://ronisbr.github.io/PrettyTables.jl/v2.4/) [@chagas2024pretty], respectively. For plots and visualizations we used both [Plots.jl](https://docs.juliaplots.org/v1.40/) [@PlotsJL] and [Makie.jl](https://docs.makie.org/v0.22/) [@danisch2021makie], in particular [AlgebraOfGraphics.jl](https://aog.makie.org/v0.9.3/). To distribute computational tasks across multiple processors, we have relied on [MPI.jl](https://juliaparallel.org/MPI.jl/v0.20/) [@byrne2021mpi].

## Reproducibility

We have taken care to set random seeds for reproducibility.