### 2025-07-23 README.qmd
adjustments to README

---
format:
  commonmark:
    variant: -raw_html+tex_math_dollars
    wrap: none
    mermaid-format: png
crossref:
  fig-prefix: Figure
  tbl-prefix: Table
output: asis
---

# CounterfactualTraining.jl

This repository contains all the code used to generate the experiments for the paper *Counterfactual Training: Teaching Models Plausible and Actionable Explanations*.

## Structure

The repository is structured like a standard Julia package with the main code base in `src/`, the documentation in `docs/` and the test suite in `test/`^[The latter two are both still incomplete]. Everything relating specifically to the paper lives under `paper/`.

### Code Base in `src/`

The main code base for the CounterfactualTraining.jl package in `src/` is relatively simple and small. It implements the core functionality for counterfactual training. The code is commented with reference to relevant parts in the paper.

### Code in `paper/`

The majority of code scripts live in `paper/experiments/`, which is a self-sufficient Julia project that depends on CounterfactualTraining.jl in the root folder. The code scripts in `paper/experiments` include Julia and bash scripts used to implement and run the experiments specific to this paper. The project depends on another self-contained helper package that lives in `paper/experiments/CTExperiments/`. This helper package makes custom functions, structs and other definitions reusable in the project (`paper/experiments/`). All package dependencies are automatically handled in scripts using Julia's package manager, Pkg.jl.

## Reproducibility

If you want to reproduce the results presented in our paper and the supplementary appendix, please see the [paper/experiments/README](/paper/experiments/README.md) for details.