### 2025-07-23 paper/experiments/README.qmd

---
format:
  commonmark:
    variant: -raw_html+tex_math_dollars
    wrap: none
    mermaid-format: png
crossref:
  fig-prefix: Figure
  tbl-prefix: Table
output: asis
engine: julia
julia: 
  exeflags: ["--project=./"]
execute:
  eval: false
  echo: true
---

# Experiments

This is the folder that contains all code used to generate the results in the paper. You can also use it to reproduce the results.

## Getting Started

To get started activate the environment in the Julia REPL:

```{julia}
# eval: false

using Pkg; Pkg.activate("paper/experiments")
```

Next, we load the `CTExperiments` package---a helper package that ships some additional functionality relevant only to the experiments run for the paper.

```{julia}
using CTExperiments
```

## Command Line Usage

```
sbatch --ntasks=$(determine_resources) --cpus-per-task=$NTHREADS myfile.sh
```

```
sbatch --ntasks=$(determine_resources --data=mydata --model=mymodel) --cpus-per-task=$NTHREADS myfile.sh --data=mydata --model=mymodel
```

:::{.callout-warning}
The extra arguments `--data=mydata --model=mymodel` should be consistent, i.e. they should be passed consistently to `determine_resources` and the whole `sbatch` command. 
:::

## Interactive Usage

### Running a single experiment

The `CTExperiments` package ships with a helper function that can be used to set up the default experiment configuration:

```{julia}
config_file = generate_template()
```

An experiment can be generated directly from that configuration file:

```{julia}
experiment = Experiment(config_file)
```