### 2025-07-24 paper/sections/other/notation.qmd

# Notation {.appendix}

## Variables and Parameters

Below we provide an overview of some notation used frequently throughout the paper:

- $\mathcal{Y}$: The output domain.
- $y^+$: The target class and also the index of the target class.
- $y^-$: The non-target class and also the index of non-the target class.
- $\mathcal{X}$: The input domain.
- $\mathbf{x}$: a single training sample.
- $\mathbf{x}^\prime$: a counterfactual.
- $t=1,...,T$: Step indicator for counterfactual search iterations.
- $\mathbf{x}^\prime_{t \leq T}$: a nascent counterfactual, defined as a counterfactual that has not yet converged.
- $\mathbf{x}^\prime_{T}$: a mature counterfactual, defined as a counterfactual that has either passed a threshold probability or exhausted all $T$ steps.
- $\mathbf{x}^+$: a training sample in the target class (ground-truth).
- $\mathbf{y}^+$: The one-hot encoded output vector for the target class.
- $\theta$: Model parameters (unspecified).
- $\Theta$: Matrix of parameters. 
- $\mathbf{M}(\cdot)$: linear predictions (logits) of the classifier.

## Formulas

### Maximum Mean Discrepancy

Maximum mean discrepancy is defined as follows,

$$
\begin{aligned}
\text{MMD}({X}^\prime,\tilde{X}^\prime) &= \frac{1}{m(m-1)}\sum_{i=1}^m\sum_{j\neq i}^m k(x_i,x_j) \\ &+ \frac{1}{n(n-1)}\sum_{i=1}^n\sum_{j\neq i}^n k(\tilde{x}_i,\tilde{x}_j) \\ &- \frac{2}{mn}\sum_{i=1}^m\sum_{j=1}^n k(x_i,\tilde{x}_j)
\end{aligned}
$$ {#eq-mmd}

where $k(\cdot,\cdot)$ is a kernel function [@gretton2012kernel]. We make use of a Gaussian kernel with a constant length-scale parameter of $0.5$. In our implementation, @eq-mmd is by default applied to the entire subset of the training data for which $y=y^+$.

## Other Conventions

In some place of this appendix, we use the terms *full/Full* (i.e. the full CT objective) and *vanilla/Vanilla* (i.e. vanilla training objective) to refer to models trained with counterfactual training (*CT*) and the baseline (*BL*), respectively.