### 2025-07-29 paper/sections/conclusion.qmd
final pass-through done

# Conclusion {#sec-conclusion}

State-of-the-art machine learning models are prone to learning complex representations that cannot be interpreted by humans. Existing work on counterfactual explanations has largely focused on designing tools to generate plausible and actionable explanations for any model. In this work, we instead hold models accountable for delivering such explanations. We introduce counterfactual training: a novel training regime that integrates recent advances in contrastive learning, adversarial robustness, and CE to incentivize highly explainable models. Through theoretical results and extensive experiments, we demonstrate that CT satisfies this goal while promoting adversarial robustness of models. Explanations generated from CT-based models are both more plausible (compliant with the underlying data-generating process) and more actionable (compliant with user-specified mutability constraints), and thus meaningful to their recipients. In turn, our work highlights the value of simultaneously improving models and their explanations.